### 2021-02-12 chill-yield-vignette.Rmd
First stab with the pasitR tools (we used these to estimates for chill and yield, different topic but similar problem)

---
title: "GHG and milk yield"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

<!-- Print pdf and word versions -->
<!-- rmarkdown::render("chill-yield-vignette.Rmd", output_format = "pdf_document") -->
<!-- rmarkdown::render("chill-yield-vignette.Rmd", output_format = "word_document") -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# devtools::install_github("hortibonn/pasitR")
library(pasitR)

# Add Unit Tests with https://testthat.r-lib.org/
# Install the released version from CRAN
# install.packages("testthat")

#The easiest way to get started is with usethis. Assuming you’re in a package directory, just run usethis::use_test("name") to create a test file, and set up all the other infrastructure you need. If you’re using RStudio, press Cmd/Ctrl + Shift + T (or run devtools::test() if not) to run all the tests in a package..

```

## Load data

```{r load_data}
ghg_variables<- read.csv("MC results/mcSimulationResults.csv")

# select the appropriate variables from the MC outputs names(ghg_variables)

#assuming 'on-farm' is the annual GHG 
chill <- ghg_variables$on_farm
#assuming 'kg_FPCM_year-farm' is the annual milk yield
yield <- ghg_variables$kg_FPCM_year

```

## Scatter plot

Using `chillscatter()` create a chill portions (x) and yield (y) scatter plot with associated and estimated densities with loess smooth linear fits density curves. Plot made with the scatter.hist() function in the plyr() package (Wickham 2011) in R (R Core Team 2017).

```{r, fig.width=7, fig.height=7}
pasitR::chillscatter(chill, yield)
```

## Kernel plot

Use `chillkernel()` or another type of graphical representation of the same relationship. `chillkernel()` performs a two-dimensional kernel density estimation for yield and chill using the `kde2d()` function in the `MASS()` package (Venables and Ripley 2002). The function produces a matrix of the estimated density (z) of yield (y) and chill (x). As the density function restricts the shape of the kernel to a bivariate normal kernel, it looks slightly different compared to the scatter plot estimates above.

```{r, fig.width=7, fig.height=7}
pasitR::chillkernel(chill, yield)
```

`chillkernel()` shows a density surface plot of chill portions (x) and yield (y). The legend shows the value for the estimated density (z). The plot is made with the `filled.contour()` function of the `graphics()` package.

In `chillkernel()` the density (z) over the entire plot integrates to one, and therefore represents the relative probability of an observation (yield along y-axis) given a specific chill portion (along x-axis).

## Estimated yield given the expected chill

`chillkernelslice()` calculates the estimated yield given the expected chill, based on a slice of 'z' from the Kernel density calcualted with `chillkernel()`. The `expectedchill` parameter is set to 30.

```{r, fig.width=7, fig.height=7}
pasitR::chillkernelslice(chill, yield, expectedchill = 30)
```

`chillkernelslice()` plots the probabilities (shown along the y-axis) for the expected yield (shown along the x-axis). Since this is a cut through the density kernel `chillkernel()`, which integrates to 1, the probability values are relative, not absolute measures.

## Chill portion intervals

`chillviolin()` determines different possible chill portion intervals by calculating the optimal interval width for chill portions using the `IQR()` function in the `stats()` package, after the Freedman-Diaconis rule (IQR = interquartile range).
`Optimal interval width for our sample = 2 * interquartile range for our sample / (total number of observations for the interquartile range for our sample)^(1/3)`

```{r, fig.width=7, fig.height=7}
pasitR::chillviolin(chill, yield)
```

`chillviolin()` shows violin plots of chill portions (x) and yield (y) with six different intervals of chill portions. Plot made with `ggplot2()` (Wickham 2016).

## Probability of yield given chill

`chillkernelslicerange()` The chill portion intervals, optimized interquartile ranges shown in `chillviolin()` can be used to select a range to slice from the density kernel `chillkernel()` as was doen for a single chill value in `chillkernelslice()`. Here we set the maximum chill to 20 and the minimum to 10.

```{r, fig.width=7, fig.height=7}

# min_chill	
# is a value of the lowest expected amount of chill for which the yield should be estimated (must be < expectedhighchill).

# 
# max_chill	
# is a value of the highest expected amount of chill for which the yield should be estimated (must be > expectedlowchill).

pasitR::chillkernelslicerange(chill, yield, min_chill = min(chill), max_chill = max(chill))
```

`chillkernelslicerange()` plots the probabilities (shown along the y-axis) for the expected yield (shown along the x-axis). As with `chillkernelslice()` the probability values shown are relative, not absolute measures. They are the result of cuts through the density kernel `chillkernel()`, which integrates to 1.

## References

R Core Team (2017) R: A Language and Environment for Statistical Computing [R Version 3.4.1 (2017-06-30) “Single Candle”] 3.4.1.

Venables, W. N., and B. D. Ripley (2002) Modern Applied Statistics with S. Fourth Edition. New York: Springer.

Wickham, Hadley (2011) Plyr: Tools for Splitting, Applying and Combining Data, Version 1.4. Available Online at Cran. r-Project. Elegant Graphics for Data Analysis. Springer.