### 2021-08-19 210617_GHG_model.Rmd
first running model (needs quality checks)

---
title: "Greenhouse gas emissions model"
author: "Cory and Eike"
bibliography:
 - references/packages.bib
 - references/papers.bib
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = F, include = F}
#install and load packages
library(decisionSupport)
library(ggplot2)
library(plyr)
library(rmarkdown)
# devtools::install_github("CWWhitney/uncertainty")
library(uncertainty)

#Automatically write R package citation entries to a .bib file
knitr::write_bib(c(.packages(), 
                   'decisionSupport'), 'references/packages.bib')

```

## Generating the GHG emissions model

We're looking at the equation put together by Joshua Wafula to check if this can easily be used by decisionSupport [@R-decisionSupport], and to possibly optimize the model.

The formulas come from Chapter 10 of the *2006 IPCC Guidelines for National Greenhouse Gas Inventories* [@dong_ipcc_2006]. The document can be found on [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/2006gl/pdf/4_Volume4/V4_10_Ch10_Livestock.pdf). The chapter is part of the book *Good Practice Guidance and Uncertainty Management in National Greenhouse Gas Inventories* [@ipcc_good_2021] , also on the [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/gp/english/). Many variables are based on the work done in *Variation in the carbon footprint of milk production on smallholder dairy farms in central Kenya* [@wilkes_variation_2020] and in *Methods and guidance to support MRV of livestock emissions* [@wilkes_methods_2019].

**Here are our changes and additions to the original model:**

```{r load_data}

library(decisionSupport)
###CW Note ####
# #### remove all the 'C:/Users/jwafula/Dropbox/UNIQUE_CCAFS/analysis/data/' folder locations throughout - 
#### these are not useful with RMD and RStudio / Git integrations
###CW Note ####
### # This project will be stored in relative file paths (i.e. same or lower folders)
input_table <- "data/livestock_ghg_input_table.csv"
```

Here we generate a `make_variables` function for testing the model function 'line by line'. We use it to take a single random sample of the provided estimates.

```{r make_variables}
###CW Note ####
### We will need to clean up the warnings later
### Reasonable ranges still missing for many variables
make_variables <- function(est,n=1){ x <- random(rho=est, n=n)
for(i in colnames(x)) assign(i, as.numeric(x[1,i]),envir=.GlobalEnv)}

# run the function on the input_table
make_variables(estimate_read_csv(input_table))
```

```{r ghg_emissions_function}
# Generate the GHG emissions function 
ghg_emissions<-function(x, varnames){

# Enteric fermentation emissions ($CH_4$)
  ####CW Note ####
  #####in these code sections I am missing the overview. 
  ##### it is not immediately clear what this is meant to do... 
  ##### from a specific publication?

# This equation is documented in "Good Practice Guidance and Uncertainty Management in National Greenhouse Gas Inventories" p. 313
#####EL CW Note ####
## We note that these equations are full of formatting mistakes that don't exactly instill confidence that all the numbers are correct.

# We start with the net energy for maintenance (NE_m; in MJ/day)
# This includes a maintenance coefficient, for which different values are recommended depending on animal type
NE_m<-maintenaince_coefficient_Cfi*(live_weight_LW^0.75)

# The energy use for specific activities is calculated by multiplying NE_m by an activity coefficient (C_activity).

NE_activity<-C_activity*NE_m

#Net energy for growth (NE_growth; MJ/day):
# here, C is an animal type coefficient
# ### CW Note #### Assuming 'C' is C_activity or should it have been animal_type
C <- C_activity
# C <- animal_type
NE_growth<-4.18*((0.0635*478/(C*FSBW))^0.75*
                   (0.96*SWG)^1.097)

# FSBW: final shrunk body weight in kg - this is the actual final body weight multiplied by 0.96 (to subtract the last meal)

##### EL CW Note ####
# SWG is the "Shrunk Weight Gain" in kg per day - we're not sure what exactly this means.

##### EL CW Note #### 
#This used to be the following equation, which we didn't find a source for:
#NE_growth<-22.02*(((live_weight_LW/(growth_coefficient_C*mature_weight_MW))^0.75)*(weight_gain_WG^1.097))

##### EL CW Note #### 
#The code doesn't include net energy mobilization by animals that lose weight (p. 314)

#Net energy for lactation

NE_lactation<-milk_yield*(1.47+0.40*milk_fat)

#variables:
#milk_yield (kg milk per day)
#milk_fat (percent)

#Draft animal services are not accounted for

#Energy need during pregnancy 
NE_pregnancy<-0.10*NE_m

##### EL CW Note #### 
#in Joshua's equation the 0.10 was a variable called pregnancy_coefficient_Cp with a value around 0.03. We're not sure where this came from.

#We've now covered all equations up to eq 10 (p. 315)

##### EL CW Note #### 
#Equation 11 contains calculation of the metabolizable energy, but that doesn't seem to be used anywhere

#Now we calculate the ratio between net energy consumed and the digestible energy consumed, for maintenance (NE_DE) and growth (NEg_DE) (eqs. 12 and 13)

NE_DE<-(1.123-(4.092*(10^-3)*dig_energy)+
            (1.126*(10^-5)*(dig_energy^2))-
              (25.4/dig_energy))

NEg_DE<-(1.164-(5.160*(10^-3)*dig_energy)+
            (1.308*(10^-5)*(dig_energy^2))-
              (37.4/dig_energy))

##### EL CW Note #### 
#the following equation usually contains the energy mobilized by animals losing weight. This wasn't calculated, so we set this to 0
NE_mobilized<-0
#same for extra energy spent by working animals
NE_w<-0

gross_energy_intake<-(((NE_m+NE_mobilized+NE_activity+NE_lactation+NE_w+NE_pregnancy)/NE_DE)+
                          (NE_growth/NEg_DE))/(dig_energy/100)

enteric_CH4<-(gross_energy_intake*Y_m*365)/55.65

##### EL CW Note #### 
######Y_m  here is the methane conversion rate -
###### Joshua used the IPCC default value of 6.5%. Full story is on page 302 (a bit more complicated, with differentiation by animal type)

# Compute the CO2 equation for enteric methane
##### EL CW Note #### 
##### #This equation originally included multiplication by 365 - this seems to have been erroneous (already done in the previous line)
#This converts methane emissions to their Global Warming Potential (CO2-equivalent of CH4 is 25)
enteric_CO2<-enteric_CH4*25

# Manure management methane ($CH_4$) and nitrous oxide ($N_2O$)
# daily volatile solid excreted (page 10.42 of 2006 IPCC Guidelines for National Greenhouse Gas Inventories)

volsol<-(gross_energy_intake*
           (1-(dig_energy/100))+
           (urinary_energy*gross_energy_intake))*
             ((1-ash_content)/18.45)

# The maximum methane emission from manure (Bo) depends on the animal type. In the input table, we
# treat these as constants for each animal type. The following line assigns the corresponding Bo value
##### EL CW Note #### 
##### # to the Bo variable. In the dataset we're using for Kenya, only two values for Bo are given: 0.17 for
# animal types 4-6 and 0.135 for animal types 1-3 and 7-12.

Bo<-eval(parse(text=paste0("Bo_animaltype_",animal_type)))

# decisions on manure management
# the management options we're considering here are the ones from the Excel table. The IPCC guidelines
# list more and partly different ones.

# We're deciding on manure management according to a decision tree (Cory will build a figure):

# question 1: how much of the manure is centrally collected? share_centrally
# question 2: of manure that's not centrally collected, how much is burned? share_central_burn
mms_burn<-(1-share_centrally)*share_central_burn
mms_pasture<-(1-share_centrally)*(1-share_central_burn)

# question 3: of centrally collected manure, how much is spread daily on cropland or pasture? share_spread
mms_spread<-share_centrally*share_spread
not_spread<-share_centrally*(1-share_spread)
# question 4: how much of the manure that's not spread is sold? share_sold
mms_sold<-not_spread*share_sold
not_sold<-not_spread*(1-share_sold)
# question 5: how much of the stored manure that's not sold is stored in liquid form? share_liquid
liquid<-not_sold*share_liquid
solid<-not_sold*(1-share_liquid)
# question 6: how much of the liquid manure is converted to biogas? share_biogas
mms_biogas<-liquid*share_biogas
mms_liquid_slurry<-liquid*(1-share_biogas)
# question 7: how much of the solid manure is composted? share_compost
mms_composted<-solid*share_compost
remain_solid<-solid*(1-share_compost)
# question 8: how much of the solid manure is stored in dry lots? share_drylot
mms_drylot<-remain_solid*share_drylot
mms_solidstore<-remain_solid*(1-share_drylot)

# the following is page 10.41 of the 2006 IPCC Guidelines for National Greenhouse Gas Inventories
mm_ch4<-(volsol*365)*
        (Bo*0.67*
           ((mms_pasture*MCFpasture)+
            (mms_spread*MCFspread)+
            (mms_drylot*MCFdrylot)+
            (mms_solidstore*MCFsolidstore)+
            (mms_composted*MCFcomposted)+
            (mms_liquid_slurry*MCFliquid)+
            (mms_biogas*MCFbiogas)+
            (mms_burn*MCFburn)+
              # the original code was lacking the 'sold' fraction - we added this for completeness
            (mms_sold*MCFsold)))

# Estimate CO2eq for manure methane
### ##### EL CW Note #### 
### this one also contained a multiplication by 365 that seemed erroneous
mm_CO2<-mm_ch4*25

# Calculate Nitrous oxide emissions from manure and urine on pasture

### ####
n_intake<-(gross_energy_intake/18.45)*((perc_crude_protein_diet/100)/6.25) # equation 10.32

# There was some code here on N retention in milk. That's apparently from equation 10.33, but it only had the first part of the equation, missing the part on weight gain. We decided to go with the alternative way of estimating N retention (figures given in Table 10.20)
# n_retention<-((milk_yield*(3.3/100))/6.38)

# Table 10.20 contains N retention coefficients:
# Dairy cows: 0.20 N_retention_dairy_cow
# Other cattle: 0.07 N_retention_other_cattle

animal_type_dairy_cow=1 ###ASSUMPTION

if(animal_type==animal_type_dairy_cow)
  n_excretion<-n_intake*(1-N_retention_dairy_cow)*365 else
    n_excretion<-n_intake*(1-N_retention_other_cattle)*365 #equation 10.31

# Direct nitrous oxide emissions (Eq 10.25 from IPCC guidelines)
mm_direct_n2o<-n_excretion*((mms_pasture*mndec_pasture)+
                           (mms_spread*mndec_spread)+
                           (mms_drylot*mndec_drylot)+
                           (mms_solidstore*mndec_solidstore)+
                           (mms_composted*mndec_composted)+
                           (mms_liquid*mndec_liquid)+
                           (mms_biogas*mndec_biogas)+
                           (mms_burn*mndec_burn)+
                           (mms_sold*mndec_sold))*(44/28)

# Nitrous oxide (N2O) from volatization (Eq 10.26 from IPCC guidelines)
mm_vol_n2o<-n_excretion*((mms_pasture*mnvc_pasture)+
                           (mms_spread*mnvc_spread)+
                           (mms_drylot*mnvc_drylot)+
                           (mms_solidstore*mnvc_solidstore)+
                           (mms_composted*mnvc_composted)+
                           (mms_liquid*mnvc_liquid)+
                           (mms_biogas*mnvc_biogas)+
                           (mms_burn*mnvc_burn)+
                           (mms_sold*mnvc_sold))*(44/28)

# Nitrous oxide (N2O) from leaching
# comment: this is probably wrong, since leached nitrogen doesn't emit (much) nitrous oxide
#mm_leach_n2o<-n_excretion*((mms_pasture*0.3)+(mms_spread*0.3)+(mms_drylot*0.3)+(mms_solidstore*0.3)+
#              (mms_composted*0.3)+(mms_liquid*0.3)+(mms_biogas*0)+(mms_burn*0))*0.0075*(44/28)

# Estimate CO2eq for manure nitrous
mm_CO2_N2Od<-mm_direct_n2o*298*365 # (298 is the CO2-equivalent of N2O)
mm_CO2_N2Ovol<-mm_vol_n2o*298*365
# mm_CO2_N2Oleach<-mm_leach_n2o*298*365
mm_CO2_N2O<-mm_CO2_N2Od+mm_CO2_N2Ovol #+mm_CO2_N2Oleach

# Emissions from feed production and transport. 
# comment: the numbers in the table seem spuriously precise.
feed_CO2<-feed_prod_CO2*feed_trans_CO2

# Total emissions
on_farm<-enteric_CH4+enteric_CO2+mm_CO2+mm_CO2_N2O+feed_CO2

#CW Note ####
# Not clear what this part does (through to the ggplot)

# Relationship between greenhouse gas intensity and milk yields. 
# Fat adjusts
#ann_adj <-kg_FPCM_year*365 
#df <- data.frame(
#  GHGi = on_farm/ann_adj, 
#  FPCM = ann_adj)

# ### ##### CW Note
# Removing this plot
# plot

#library(ggplot2)
#ggplot(df, aes(x = FPCM, y = GHGi)) +geom_point() + ylim(0,50)

return(list(enteric_CH4=enteric_CH4,
            #CW NOTE ####
            # ####working backwards from 'enteric_CO2' I see that 
            # #### this is the product of something twice multiplied by 365
            # ####### 
            milk_yield=milk_yield,
            enteric_CO2=enteric_CO2,
            mm_CO2=mm_CO2,
            mm_CO2_N2O=mm_CO2_N2O,
            feed_CO2=feed_CO2,
            on_farm=on_farm))
}

```

Apply the `decisionSupport` tools for running the Monte Carlo [@R-decisionSupport]. See the vignette guides on *applying the mcSimulation function* [@fernandez_applying_2021] and *simulating the cost effectiveness of controlled burning* [@whitney_simulating_2017] for examples.

```{r monte_carlo_simulation}

GHG_simulation <- mcSimulation(
  estimate = estimate_read_csv("data/livestock_ghg_input_table.csv"),
  model_function = ghg_emissions,
  numberOfModelRuns = 1e2, #100 for now
  functionSyntax = "plainNames"
)

```

```{r plot_distributions}

plot_distributions(mcSimulation_object = GHG_simulation,
                   vars = "enteric_CH4",
                   method = "boxplot_density",
                   old_names = "enteric_CH4",
                   new_names = "Methane Emissions from Enteric Fermentation")

```

# uncertainty

Here we apply a collection of functions for use in assessing uncertainty from the `uncertainty` package. Install using `devtools::install_github("CWWhitney/uncertainty")`.

Here we use the `varscatter()` function to create a scatter plot of an influence variable of interest `in_var` (x) and an outcome variable `out_var` (y) with associated and estimated densities with loess smooth linear fits density curves. Plot made with the scatter.hist() function in the plyr() package [@R-plyr] in R [@R-base].

Check the `milk_yield` effect on `enteric_CH4`.

```{r uncertainty_package}
uncertainty::varscatter(in_var = GHG_simulation$x$milk_yield, out_var = GHG_simulation$y$enteric_CH4)
```

Use `varkernel()` for another type of graphical representation of the same relationship. `varkernel()` performs a two-dimensional kernel density estimation for the expected value of the outcome variable `out_var` and the influence variable of interest `in_var` using the `kde2d()` function in the `MASS()` package (Venables and Ripley 2002). The function produces a matrix of the estimated density (z) of the expected value of the outcome variable `out_var` (y) and the influence variable of interest `in_var` (x). It looks slightly different than the scatter plot estimates above because the density function restricts the shape of the kernel to a bivariate normal kernel.

Check the effect of `in_var` `milk_yield` on the `out_var` `enteric_CH4`.

```{r name_in_out_vars}
in_var  <- GHG_simulation$x$milk_yield
out_var <- GHG_simulation$y$enteric_CH4
```

```{r varkernel}
uncertainty::varkernel(in_var = in_var, out_var = out_var)
```

`varkernel()` shows a density surface plot of a given influence variable of interest `in_var` (x) and the outcome variable `out_var` (y). The legend shows the value for the estimated density (z). The plot is made with the `filled.contour()` function of the `graphics()` package.

In `varkernel()` the density (z) over the entire plot integrates to one, and therefore represents the relative probability of an observation (the outcome variable `out_var` along y-axis) given a specific value for a given influence variable of interest `in_var` (along x-axis).

## Estimated outcome variable `out_var` given the expected variable

`varkernelslice()` calculates the estimated outcome variable `out_var` given the expected value of the influence variable of interest `in_var`, based on a slice of 'z' from the Kernel density calculated with `varkernel()`. The `expectedvariable` parameter is set to 30.

```{r varkernelslice}
uncertainty::varkernelslice(in_var = in_var, out_var = out_var, expectedin_var = 30)
```

`varkernelslice()` plots the probabilities (shown along the y-axis) for the expected value of the outcome variable `out_var` (shown along the x-axis). Since this is a cut through the density kernel `varkernel()`, which integrates to 1, the probability values are relative, not absolute measures.

## Influencing variable intervals

`varviolin()` determines different possible influencing variable intervals `in_var` by calculating the optimal interval width for the given variable values. This is done using the `IQR()` function in the `stats()` package, after the Freedman-Diaconis rule (IQR = interquartile range).

$Os = 2 * IQRs / (obs_IQRs)^{(1/3)}$

Where $Os$ is The optimal interval width for our sample, $IQRs$ is the interquartile range for our sample and $obs_IQRs$ is the total number of observations for the interquartile range for our sample.

```{r varviolin}
uncertainty::varviolin(in_var = in_var, out_var = out_var)
```

`varviolin()` shows violin plots of influencing variable `in_var` values (x) and outcome variable `out_var` (y) values, with six different intervals of influencing variable `in_var` values. Plot made with `ggplot2()` [@R-ggplot2].

## Probability of outcome variable given variable

`varkernelslicerange()` shows the influencing variable `in_var` value intervals, optimized interquartile ranges shown in `varviolin()` can be used to select a range to slice from the density kernel `varkernel()` as was done for a single variable value in `varkernelslice()`. Here we set the maximum expected value for `in_var` to 4 and the minimum expected value to 1.

```{r varkernelslicerange}
uncertainty::varkernelslicerange(in_var = in_var, out_var = out_var, 
                                 min_in_var = 1, max_in_var = 4)
```

`varkernelslicerange()` plots the probabilities (shown along the y-axis) for the expected value of the outcome variable `out_var` (shown along the x-axis). As with `varkernelslice()` the probability values shown are relative, not absolute measures. They are the result of cuts through the density kernel `varkernel()`, which integrates to 1.

This document was generated using R Markdown [@R-rmarkdown].

# References