### 2022-06-14 Manuscript_figures.Rmd
first complete draft of methods and results

---
title: "Accounting for uncertainty in simulating greenhouse gas emissions from dairy cattle in Kenya"
author: "Eike Luedeling, Cory Whitney, Andreas Wilkes, Todd Rosenstock"
bibliography:
 - references/packages.bib
 - references/papers.bib
output:
  html_document:
    theme: united
    toc: true
    toc_float: true
    number_sections: true
    code_folding: hide
---

# Setting up

Note that the code for all operations is hidden, so that it doesn't clutter the document. You can make it appear by clicking the code button on the right.

We first need to install and load all required packages.

```{r setup, warning=FALSE, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

#install and load packages
# devtools::install_github("CWWhitney/uncertainty")
library(DiagrammeR)
library(decisionSupport)
library(ggplot2)
library(kableExtra)
library(magrittr)
library(plyr)
library(rmarkdown)
library(uncertainty)
library(chillR)
library(raster)
library(reshape2)
library(patchwork)

knitr::write_bib(c(.packages(), 
                   'DiagrammeR',
                   'decisionSupport',
                   'ggplot2',
                   'kableExtra',
                   'magrittr',
                   'plyr',
                   'rmarkdown',
                   'uncertainty',
                   'chillR',
                   'raster',
                   'reshape2',
                   'patchwork'), 'references/packages.bib')
#resolves an issue with the scientific style of numbers,answer from @Paul Hiemstra: https://stackoverflow.com/a/25947542/4061993 
#from the documentation of ?options:
options(scipen=999)
```

```{r, warning=FALSE, eval=FALSE}
devtools::install_github("CWWhitney/uncertainty")
library(DiagrammeR)
library(decisionSupport)
library(ggplot2)
library(kableExtra)
library(magrittr)
library(plyr)
library(rmarkdown)
library(uncertainty)
library(chillR)
library(raster)
library(reshape2)
library(patchwork)
```

## Load emission calculation function

Greenhouse gas emissions from livestock are calculated according to Chapter 10 of the *2006 IPCC Guidelines for National Greenhouse Gas Inventories* [@dong_ipcc_2006] (Volume 4 - Agriculture, Forestry and Other Land Use). The document can be found on [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/2006gl/pdf/4_Volume4/V4_10_Ch10_Livestock.pdf). The chapter is part of the book *Good Practice Guidance and Uncertainty Management in National Greenhouse Gas Inventories* [@ipcc_good_2021], also on the [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/gp/english/). Many variables are based on the work done in *Variation in the carbon footprint of milk production on smallholder dairy farms in central Kenya* [@wilkes_variation_2020] and in *Methods and guidance to support MRV of livestock emissions* [@wilkes_methods_2019]. We also draw on recommendations provided in Chapter 10 of Volume 4 of the *2019 Refinement to the 2006 IPCC Guidelines for National Greenhouse Gas Inventories* [@ipcc_2019], which is also accessible via [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/2019rf/vol4.html).

The equations provided in these documents, in particular in the 2006 guideline document [@dong_ipcc_2006] were converted into an R function called `ghg_emissions`, which is defined here (hit the `Code` button on the right).

```{r ghg_emissions_function}
# Generate the GHG emissions function 
ghg_emissions<-function(x, varnames){

  # Enteric fermentation emissions ($CH_4$)

  # The main source of all this is Chapter 10 of the *2006 IPCC Guidelines 
  # for National Greenhouse Gas Inventories* 
  # (Volume 4 - Agriculture, Forestry and Other Land Use). 
  # The document can be found on the IPCC web 
  # https://www.ipcc-nggip.iges.or.jp/public/2006gl/pdf/4_Volume4/V4_10_Ch10_Livestock.pdf

  # We start with the net energy for maintenance (NE_m; in MJ/day)
  # This includes a maintenance coefficient, 
  # for which different values are recommended depending on animal type

  # The following animal types were used in the example:

  # animal type 1 - 'intact' grown-up males >=36 months
  # animal type 2 - castrated grown-up males >=36 months
  # animal type 3 - males 12-36 months
  # animal type 4 - lactating cows
  # animal type 5 - lactating cows
  # animal type 6 - lactating cows
  # animal type 7 - weaned young females <= 12 months
  # animal type 8 - weaned young males <= 12 months
  # animal type 9 - females 12-36 months
  # animal type 10 - pre-weaned female calves (<12 months)
  # animal type 11 - pre-weaned male calves (<12 months)
  # animal type 12 - first-time pregnant cows

  # animal types 4, 5 and 6 are all lactating cows - 
  # important to note here that the original data table 
  # appears to have worked with the actual number of months 
  # that cows were pregnant. This resulted in a weighted mean 
  # of the various maintenance coefficients. 
  # We're not sure where we could get such data. 
  # In the original dataset, types 4, 5 and 6 
  # referred to different pregnancy/lactation states, 
  # but it doesn't seem like this scheme was rigorously followed.

  # We make placeholder vectors to collect emissions data 
  # for all animal types
  mm_N2O_CO2eq<-c()
  mm_CH4_CO2eq<-c()
  enteric_CH4_CO2eq<-c()
  total_milk<-c()

  # Now we start a long loop, 
  # where emissions are calculated for each animal type

  for (animal_type in 1:12)
  {
    # define C and Cfi based on sex (Page 10.17, below Equation 10.6)
    if(animal_type %in% c(1,3,8,11)) growth_coefficient_C<-C_intact_males # males (1.2)
    if(animal_type %in% c(2)) growth_coefficient_C<-C_castrates # castrates (1.0)
    if(animal_type %in% c(4,5,6,7,9,10,12)) growth_coefficient_C<-C_females # females (0.8)

    if(animal_type %in% c(4,5,6)) Cfi<-maintenance_coefficient_Cfi_lact_cow # lactating cows (0.386)
    if(animal_type %in% c(1)) Cfi<-maintenance_coefficient_Cfi_bulls # bulls (0.37)
    if(animal_type %in% c(2,3,7,8,9,10,11,12)) Cfi<-maintenance_coefficient_Cfi_other # others (0.322)

    # define mature body weight (assuming this needs to be disaggregated by sex, for Eqation 10.6)
    if(animal_type %in% c(1,2,3,8,11)) mature_weight_MW<-mature_weight_MW_male
    if(animal_type %in% c(4,5,6,7,9,10,12)) mature_weight_MW<-mature_weight_MW_female

    # define milk yield
    if(animal_type %in% c(1,2,3,7,8,9,10,11,12)) milk_yield<-0
    if(animal_type %in% c(4,5,6)) milk_yield<-eval(parse(text=paste0("milk_yield_",animal_type)))

    #define pregnancy coefficient (full Cp for first-time pregnant, lower for later in most cases)
    if(animal_type==12) Cp<-Cp_0
    if(animal_type==4) Cp<-Cp_0*Cp_rate_4
    if(animal_type==5) Cp<-Cp_0*Cp_rate_5
    if(animal_type==6) Cp<-Cp_0*Cp_rate_6
    if(animal_type %in% c(1,2,3,7,8,9,10,11)) Cp<-0

    N_animal_type<-eval(parse(text=paste0("N_animal_type_",animal_type)))

    # define activity coefficient Ca (Table 10.5)
    # Stall 0.00 - Animals are confined to a small area 
    #     (i.e., tethered, pen, barn) with the result that they expend 
    #     very little or no energy to acquire feed.
    # Pasture 0.17 - Animals are confined in areas with sufficient forage 
    #     requiring modest energy expense to acquire feed.
    # Grazing large areas 0.36 - Animals graze in open range land 
    #     or hilly terrain and expend significant energy to acquire feed.

    if(feeding_system==1) C_activity<-Ca_stall # Stall (0.0)
    if(feeding_system==2) C_activity<-Ca_pasture # Stall (0.17)
    if(feeding_system==3) C_activity<-Ca_large_grazing # Stall (0.36)

    # Mean live weight (kg) of animals surely varies by animal type, 
    # so we need separate estimates for them
    live_weight_LW<-eval(parse(text=paste0("live_weight_LW_",animal_type)))

    # Mean weight gain of animal type (kg/day)
    weight_gain_WG<-eval(parse(text=paste0("weight_gain_WG_",animal_type)))

    # Calculate the Net Energy for Maintenance (NE_m) (MJ/day) (Equation 10.3)
    NE_m<-Cfi*(live_weight_LW^0.75)

    # Calculate the Net Energy for Activity (NE_activity) (MJ/day) (Equation 10.4)
    NE_activity<-C_activity*NE_m

    #Net energy for growth (NE_growth) (MJ/day) (Equation 10.6)
    NE_growth<-22.02*
      ((live_weight_LW/(growth_coefficient_C*mature_weight_MW))^0.75)*
      (weight_gain_WG^1.097)

    # Net energy for lactation
    NE_lactation<-milk_yield*(1.47+0.40*milk_fat)

    #Energy need during pregnancy (Equation 10.13)
    NE_pregnancy<-Cp*NE_m

    # Ratio of net energy available in a diet for maintenance 
    # to digestible energy consumed (Equation 10.14)
    REM<-1.123-(4.092/1000*DE_perc)+(1.126/100000*DE_perc^2)-(25.4/DE_perc)

    # Ratio of net energy available for growth in a diet 
    # to digestible energy consumed (Equation 10.15)
    REG<-1.164-(5.160/1000*DE_perc)+(1.308/100000*DE_perc^2)-(37.4/DE_perc)

    # Gross Energy (GE) MJ/day
    NE_work<-0 # work animals aren't covered here, so we set this to zero
    GE<-(((NE_m+NE_activity+NE_lactation+NE_work+NE_pregnancy)/REM)+
           (NE_growth/REG))/(DE_perc/100)

    # CH4 emission factors for enteric fermentation from a livestock category 
    # (Equation 10.21)
    enteric_CH4<-(GE*Y_m*365)/55.65 # kg CH4 per animal per year

    # Compute the CO2-equivalent of methane emissions for enteric methane
    # This converts methane emissions to their Global Warming Potential 
    # (CO2-equivalent of CH4 is 25)
    enteric_CH4_CO2eq[animal_type]<-N_animal_type*enteric_CH4*25

    # Manure management methane ($CH_4$) and nitrous oxide ($N_2O$)
    # daily volatile solid excreted (Equation 10.24)

    volsol<-(GE*
               (1-(DE_perc/100))+
               (urinary_energy*GE))*
      ((1-ash_content)/18.45)

    # The maximum methane emission from manure (Bo) depends on the animal type. 
    # In the input table, we treat these as constants for each animal type. 
    # The following line assigns the corresponding Bo value

    Bo<-eval(parse(text=paste0("Bo_animaltype_",animal_type)))

    # CH4 emission factor from manure management (Equation 10.23)
    mm_ch4<-(volsol*365)*
      (Bo*0.67*
         ((mms_pasture*MCFpasture)+
            (mms_spread*MCFspread)+
            (mms_drylot*MCFdrylot)+
            (mms_solidstore*MCFsolidstore)+
            (mms_composted*MCFcomposted)+
            (mms_liquid*MCFliquid)+
            (mms_biogas*MCFbiogas)+
            (mms_burn*MCFburn)+
          # the original code was lacking the 'sold' fraction - 
          # we added this for completeness
          (mms_sold*MCFsold)))

    # Estimate CO2eq for manure methane
    mm_CH4_CO2eq[animal_type]<-N_animal_type*mm_ch4*25

    # Calculate Nitrous oxide emissions from manure and urine on pasture

    # estimate N intake (kg/animal/day) (Equation 10.32)
    N_intake<-(GE/18.45)*((perc_crude_protein_diet/100)/6.25)

    # estimate N retention (kg/animal/day) (Equation 10.33)

    if(weight_gain_WG==0)
      N_retention <- ((milk_yield*(1.9+0.4*milk_fat)/100)/6.38)

    if(!weight_gain_WG==0)
      N_retention<-((milk_yield*(1.9+0.4*milk_fat)/100)/6.38)+
      ((weight_gain_WG*(268-(7.03*NE_growth/weight_gain_WG)))/(1000*6.25))

    # estimate N excretion 
    # (kg N/animal/yr) (Equation 10.31)
    N_excretion<-N_intake*(1-N_retention)*365

    # Direct nitrous oxide emissions 
    # (kg N2O/animal/yr) (Eq 10.25 from IPCC guidelines)
    mm_direct_n2o<-N_excretion*((mms_pasture*mndec_pasture)+
                                  (mms_spread*mndec_spread)+
                                  (mms_drylot*mndec_drylot)+
                                  (mms_solidstore*mndec_solidstore)+
                                  (mms_composted*mndec_composted)+
                                  (mms_liquid*mndec_liquid)+
                                  (mms_biogas*mndec_biogas)+
                                  (mms_burn*mndec_burn)+
                                  (mms_sold*mndec_sold))*(44/28)

    # N losses from volatilization 
    # (kg N/animal/yr) (Eq 10.26 from IPCC guidelines)
    mm_vol_n<-N_excretion*((mms_pasture*mnvc_pasture)+
                             (mms_spread*mnvc_spread)+
                             (mms_drylot*mnvc_drylot)+
                             (mms_solidstore*mnvc_solidstore)+
                             (mms_composted*mnvc_composted)+
                             (mms_liquid*mnvc_liquid)+
                             (mms_biogas*mnvc_biogas)+
                             (mms_burn*mnvc_burn)+
                             (mms_sold*mnvc_sold))

    # indirect N2O emissions due to volatilization of N from manure management 
    # (kg N2O/animal/year) (Equation 10.27)
    mm_vol_n2o<-mm_vol_n*EF4*(44/28)

    # Nitrous oxide (N2O) from leaching (kg N/animal/year) (equation 10.28)
    mm_leach_n<-N_excretion*((mms_pasture*frac_leach_pasture)+
                               (mms_spread*frac_leach_spread)+
                               (mms_drylot*frac_leach_drylot)+
                               (mms_solidstore*frac_leach_solidstore)+
                               (mms_composted*frac_leach_composted)+
                               (mms_liquid*frac_leach_liquid)+
                               (mms_biogas*frac_leach_biogas)+
                               (mms_burn*frac_leach_burn)+
                               (mms_sold*frac_leach_sold))

    # indirect N2O emissions due to leaching of N from manure management 
    # (kg N2O/animal/year) (Equation 10.29)
    mm_leach_n2o<-mm_leach_n*EF5*(44/28)

    # CW EL Comment ####
    # # comment: we removed this, it was probably wrong, 
    # since leached nitrogen doesn't emit (much) nitrous oxide
    #mm_leach_n2o<-n_excretion*((mms_pasture*0.3)+(mms_spread*0.3)+(mms_drylot*0.3)+(mms_solidstore*0.3)+
    #              (mms_composted*0.3)+(mms_liquid*0.3)+(mms_biogas*0)+(mms_burn*0))*0.0075*(44/28)

    # Estimate CO2eq for manure nitrous
    mm_N2O_CO2eq[animal_type]<-N_animal_type*(mm_direct_n2o+mm_vol_n2o+mm_leach_n2o)*298 # (298 is the CO2-equivalent of N2O)

    total_milk[animal_type]<-N_animal_type*milk_yield*365

  }

  # Emissions from feed production and transport.

  feed_CO2<-feed_prod_CO2+feed_trans_CO2

  # Total emissions

  on_farm<-sum(enteric_CH4_CO2eq+mm_CH4_CO2eq+mm_N2O_CO2eq)+feed_CO2

  total_animals<-sum(sapply(1:12,function(x) eval(parse(text=paste0("N_animal_type_",x)))))

  return(list(enteric_CH4=enteric_CH4,
              milk_yield=sum(total_milk),
              enteric_CH4_CO2eq=sum(enteric_CH4_CO2eq),
              mm_CH4_CO2eq=sum(mm_CH4_CO2eq),
              mm_N2O_CO2eq=sum(mm_N2O_CO2eq),
              feed_CO2=feed_CO2,
              on_farm=on_farm,
              pc_enteric_CH4=enteric_CH4/total_animals,
              pc_milk_yield=sum(total_milk)/total_animals,
              pc_enteric_CH4_CO2eq=sum(enteric_CH4_CO2eq)/total_animals,
              pc_mm_CH4_CO2eq=sum(mm_CH4_CO2eq)/total_animals,
              pc_mm_N2O_CO2eq=sum(mm_N2O_CO2eq)/total_animals,
              pc_feed_CO2=feed_CO2/total_animals,
              pc_on_farm=on_farm/total_animals

              ))
}

```

This equation takes a wide range of inputs. Many of these are usually not known precisely, yet the classic IPCC equation does not account for potential errors and uncertainties. Here, we express this uncertainty by defining distributions for each variable. The following table (**data/livestock_ghg_input_table.csv**) shows all the variables, as well as examples of distributions that may describe their quantities. Distributions are defined according to the distribution shape, as well as upper and lower bounds of 90% confidence intervals. We use the following distribution shapes:

* const - constants (upper and lower bounds are the same)
* posnorm - positive normal distribution; this is a normal distribution that is truncated at 0.

The decisionSupport package [@R-decisionSupport], which we use to run the simulations, also accommodates other distribution types, but we did not consider these appropriate here. The positive normal distribution (and other distributions such as normal or lognormal distributions) are defined based on the 5% and 95% quantiles. Note that it isn't always possible to fit a positive normal distribution based on such specifications, especially if the 5% quantile is set so close to zero that the remaining 5% of the data do not **fit** in the space between zero and the lower bound. In such cases, decisionSupport [@R-decisionSupport] returns a warning message, but it still does its best to generate the requested distribution. It is worth being on the lookout for the respective warning messages and try to adjust the bounds of the distribution to **possible** values.

```{r load_data, include=FALSE}

input_table <- "data/livestock_ghg_input_table.csv"

options(knitr.kable.NA = '', encoding = "Windows-1252")
 read.csv(input_table)[,c(1:2,4,7,5:6)]   %>%
    kableExtra::kbl(digits = 2)  %>%
    kableExtra::kable_classic_2(full_width = F) %>% 
 scroll_box(height = "300px")

```

# Monte Carlo simulation

With the model equation and the input table defined above, we can run a Monte Carlo simulation that returns a probability distribution for the greenhouse gas emissions resulting from the particular herd that is specified in the input table. We use the functions provided by the decisionSupport package [@R-decisionSupport] in the R programming language [@R-base].

```{r, warning=FALSE,eval=FALSE}
input_table <- "data/livestock_ghg_input_table.csv"
GHG_emissions<-mcSimulation(estimate=estimate_read_csv(input_table),
                            model_function=ghg_emissions,
                            numberOfModelRuns=1000,
                            functionSyntax="plainNames")
```

## Running a Monte Carlo simulation with scenarios

The `mcSimulation` function takes random draws for each of the input variables from the distributions specified in the input table. The distributions that are used are the same for every run of the Monte Carlo simulation. This may be desirable in homogeneous populations where all combinations of random values are plausible, but it often misses the mark where populations are composed of particular member types with specific characteristics. To account for populations that are heterogeneous, which is the case for livestock-keeping households in Kenya, we use the `scenario_mc` function, which allows specifying particular subgroups. Defining such scenarios can be desirable when it's necessary to simulate outcomes for particular farm types, climate scenarios etc. To do this, we would normally have to define separate input tables and run separate simulations. This is quite inconvenient. Here, we use a function that can do this job automatically.

The `scenario_mc` function in the `decisionSupport` package [@R-decisionSupport] consists of a wrapper around the `mcSimulation` function. The function can modify the input table according to a scenario file we provide. This file contains information on which variables should be modified for each scenario and how many model runs should be included in the simulation. The function call is essentially the same as for the `mcSimulation` function, but with an additional `scenarios` option which imports the contents of a .csv file.

The .csv file for the `scenarios` option specifies variables that should be modified for each of the scenarios, as well as the number of model runs for it. Each column beyond the first two columns in the file contains information for a particular scenario. This is restricted to the parameters that *should be changed* for this scenario (all others don't have to be included).

Here's an example `scenarios` file:

```{r}
read.table("data/scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM",header=TRUE)   %>%
    kableExtra::kbl(digits = 2)  %>%
    kableExtra::kable_classic_2(full_width = F)

```

The first row ("Runs" in the first column) indicates the number of runs. Each subsequent row indicates which value from the input table should be modified, so the string in the first column has to correspond to a variable name in the input table. The second column indicates which values should be changed, with options being "lower", "upper", "distribution" and "both". If "both" is selected, both the lower and upper bounds are modified (this should only be done for variables with a constant distribution, since the replacement will be for both the upper and lower bound).

We define subgroups based on the populations of households described by a household survey in Kenya [@wilkes_variation_2020; @wilkes_methods_2019]. Each of the farms encountered in this survey was converted into one scenario, for which Monte Carlo simulation is then used to produce multiple replicates.

## Household-based scenarios

The information for the household is extracted from a spreadsheet that contains all the survey responses [@wilkes_variation_2020; @wilkes_methods_2019].

### Probabilistic scenarios

Extracting all relevant information from the survey results spreadsheet is a slightly complex task, which is accomplished by the following code. The code includes specification of lower and upper bounds of probability distributions for most of the variables. We keep the herd composition constant, but all variables that in all likelihood weren't determined with high precision and accuracy are equipped with such uncertainty estimates. Estimates of the uncertainty for model input variables were extracted from the IPCC documents, or, if no such indications were available, based on our expert judgement.

```{r}
# import the data from the survey and ensure that the .csv encoding can be read by R

Kenyaherds<-read.table("data/Kenya_baseline_individual_cow_wNotes.csv", sep = ",",
                       fileEncoding = "UTF-8-BOM", header=TRUE)

# count the number of animals for each type

animals<-table(Kenyaherds[,c("QQNO","animal_type")])

variables_to_update_const<-c(paste0("N_animal_type_",1:12),
                             "feeding_system")

scenarios<-data.frame(Variable=variables_to_update_const,param="both")

variables_to_update_posnorm<-c(paste0("milk_yield_",4:6),
                             paste0("Cp_rate_",4:6),
                             "feed_prod_CO2",
                             "feed_trans_CO2",
                             paste0("live_weight_LW_",1:12),
                             paste0("weight_gain_WG_",1:12),
                             "DE_perc",
                             "perc_crude_protein_diet")

scenarios<-rbind(scenarios,data.frame(Variable=rep(variables_to_update_posnorm,each=3),param=c("lower","upper","distribution")))

for(i in 1:nrow(animals))
  {coln<-paste0("Farm_",row.names(animals)[i])
   scenarios[,coln]<-NA
     scenarios[which(scenarios$Variable %in% paste0("N_animal_type_",1:12)),coln]<-animals[i,]
}

# for some animal-scale variables (e.g. milk yield), we need averages per farm

### error estimates for variables
milk_yield_error<-0.275 # Migose et al. (2020, https://www.sciencedirect.com/science/article/abs/pii/S1871141319307371 )
feed_trans_error<-0.25 # Vellinga et al (2013, https://edepot.wur.nl/254098)
feed_prod_error<-0.25 # Vellinga et al (2013, https://edepot.wur.nl/254098)
live_weight_error<-0.145 # Goopy et al. (2018, https://www.publish.csiro.au/an/an16577 ) 
weight_gain_error<-0.145 # Goopy et al. (2018, https://www.publish.csiro.au/an/an16577 ) 
digestible_energy_error<-0.1
crude_protein_error<-0.1

# farm-scale variables

Farms<-unique(Kenyaherds$QQNO)

for (f in Farms)
  {coln<-paste0("Farm_",f)
  # update feeding system 
  scenarios[which(scenarios$Variable == "feeding_system"),coln]<-
     median(Kenyaherds[which(Kenyaherds$QQNO==f),"system"])

  # update digestible energy (mean for all animals for now) (with error)
  scenarios[which(scenarios$Variable == "DE_perc"),coln][1:2]<-
     mean(Kenyaherds[which(Kenyaherds$QQNO==f),"dig_energy"])*
                           c(1-digestible_energy_error,1+digestible_energy_error)

  # update crude protein (mean for all animals for now) (with error)
  scenarios[which(scenarios$Variable == "perc_crude_protein_diet"),coln][1:2]<-
     mean(Kenyaherds[which(Kenyaherds$QQNO==f),"CP."])*
                           c(1-crude_protein_error,1+crude_protein_error)

  # update feed production CO2 (with error)
   scenarios[which(scenarios$Variable == "feed_prod_CO2"),coln][1:2]<-
     sum(Kenyaherds[which(Kenyaherds$QQNO==f),"feed_kgCO2"])*
                           c(1-feed_prod_error,1+feed_prod_error)

  # update feed transport CO2 (with error)
   scenarios[which(scenarios$Variable == "feed_trans_CO2"),coln][1:2]<-
     sum(Kenyaherds[which(Kenyaherds$QQNO==f),"feed.transport_kgCO2"])*
                           c(1-feed_trans_error,1+feed_trans_error)

   # update milk yield for each cow type (animal type 4-6) (with error)
   for(typ in 4:6)
     if(sum(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ)>0)
       {scenarios[which(scenarios$Variable==paste0("milk_yield_",typ)),coln][1:2]<-
         mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
                         "milk_yield"])*c(1-milk_yield_error,1+milk_yield_error)
       scenarios[which(scenarios$Variable==paste0("CP_rate_",typ)),coln][1:2]<-
         mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
                         "Cp"]/0.1) *c(1,1)

     }
  # update live weight and weight gain for each animal type (with error)
   for(typ in 1:12)
     if(sum(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ)>0)
       {scenarios[which(scenarios$Variable==paste0("live_weight_LW_",typ)),coln][1:2]<-
         mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
                         "live_weight_LW"])*c(1-live_weight_error,1+live_weight_error)
       scenarios[which(scenarios$Variable==paste0("weight_gain_WG_",typ)),coln][1:2]<-
         mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
                         "weight_gain_WG"])*c(1-weight_gain_error,1+weight_gain_error)
     }

   for(varia in variables_to_update_posnorm)
   if(sum(is.na(scenarios[which(scenarios$Variable==varia),coln][1:2]))==0)   
     if(!equals(scenarios[which(scenarios$Variable==varia),coln][1],
               scenarios[which(scenarios$Variable==varia),coln][2]))
        scenarios[which(scenarios$Variable==varia&
                   scenarios$param=="distribution"),coln]<-"posnorm" else
            scenarios[which(scenarios$Variable==varia&
                   scenarios$param=="distribution"),coln]<-"const"

   }

write.csv(scenarios,"data/farm_scenarios.csv",row.names = FALSE)

read.table("data/farm_scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM",header=TRUE)[1:20,1:5]  %>%
    kableExtra::kbl(digits = 2)  %>%
    kableExtra::kable_classic_2(full_width = F)

```

This is the resulting file (only the first 3 farms are shown, and only the first 20 rows):

The first few rows in this table adjust the herd composition parameters, so that the scenario herd corresponds to the herd of the specific household. Other parameters including the milk yield per cow are also adjusted.

### Precise scenarios

We also make precise scenarios that contain only the exact values that were observed in the survey. This is needed for illustrating the impact of accounting for uncertainty later. For these scenarios, we follow the commonly used practice of applying default values recommended by the various IPCC documents mentioned above [@dong_ipcc_2006;@ipcc_2019;@ipcc_good_2021].

```{r}
# count the number of animals for each type

animals<-table(Kenyaherds[,c("QQNO","animal_type")])

variables_to_update_posnorm<-c(paste0("milk_yield_",4:6),
                             "feed_prod_CO2",
                             "feed_trans_CO2",
                             paste0("live_weight_LW_",1:12),
                             paste0("weight_gain_WG_",1:12),
                             "DE_perc",
                             "perc_crude_protein_diet")

precise_inputs<-read.csv(input_table)[,c(1:8)] 
precise_vars<-precise_inputs[which(!is.na(precise_inputs$IPCC_default)),]

precise_scenarios<-data.frame(Variable=c(variables_to_update_const,variables_to_update_posnorm,precise_vars$variable),param="both")

                                      #   precise_estimates),param="both")

for(i in 1:nrow(animals))
  {coln<-paste0("Farm_",row.names(animals)[i])
   precise_scenarios[,coln]<-NA
     precise_scenarios[which(precise_scenarios$Variable %in% paste0("N_animal_type_",1:12)),coln]<-animals[i,]
}

# farm-scale variables

Farms<-unique(Kenyaherds$QQNO)

for (f in Farms)
  {coln<-paste0("Farm_",f)
  # update feeding system 
  precise_scenarios[which(precise_scenarios$Variable == "feeding_system"),coln]<-
     median(Kenyaherds[which(Kenyaherds$QQNO==f),"system"])

  # update digestible energy (mean for all animals for now) (with error)
  precise_scenarios[which(precise_scenarios$Variable == "DE_perc"),coln]<-
     mean(Kenyaherds[which(Kenyaherds$QQNO==f),"dig_energy"])

  # update crude protein (mean for all animals for now) (with error)
  precise_scenarios[which(precise_scenarios$Variable == "perc_crude_protein_diet"),coln]<-
     mean(Kenyaherds[which(Kenyaherds$QQNO==f),"CP."])

  # update feed production CO2 (with error)
   precise_scenarios[which(precise_scenarios$Variable == "feed_prod_CO2"),coln]<-
     sum(Kenyaherds[which(Kenyaherds$QQNO==f),"feed_kgCO2"])

  # update feed transport CO2 (with error)
   precise_scenarios[which(precise_scenarios$Variable == "feed_trans_CO2"),coln]<-
     sum(Kenyaherds[which(Kenyaherds$QQNO==f),"feed.transport_kgCO2"])

   # update milk yield for each cow type (animal type 4-6) (with error)
   for(typ in 4:6)
     if(sum(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ)>0)
       {precise_scenarios[which(precise_scenarios$Variable==paste0("milk_yield_",typ)),coln]<-
         mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
                         "milk_yield"])
       precise_scenarios[which(precise_scenarios$Variable==paste0("CP_rate_",typ)),coln]<-
         mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
                         "Cp"]/0.1)

     }
  # update live weight and weight gain for each animal type (with error)
   for(typ in 1:12)
     if(sum(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ)>0)
       {precise_scenarios[which(precise_scenarios$Variable==paste0("live_weight_LW_",typ)),coln]<-
         mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
                         "live_weight_LW"])
       precise_scenarios[which(precise_scenarios$Variable==paste0("weight_gain_WG_",typ)),coln]<-
         mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
                         "weight_gain_WG"])
     }

   # now add the best estimates from the IPCC guidelines to the precise scenarios

   for(vv in 1:nrow(precise_vars))
     precise_scenarios[which(precise_scenarios$Variable==precise_vars$variable[vv]),coln]<-
     precise_vars$IPCC_default[vv]

    }

write.csv(precise_scenarios,"data/farm_scenarios_precise.csv",row.names = FALSE)

```

## Running the simulation

Now we can run the Monte Carlo simulation with these scenarios.

```{r monte_carlo_simulation, warning=FALSE, eval=FALSE}

GHG_simulation_scenarios<-scenario_mc(base_estimate = estimate_read_csv("data/livestock_ghg_input_table.csv"),
                                 scenarios = read.csv("data/farm_scenarios.csv",
                                                      fileEncoding = "UTF-8-BOM"),
                                 model_function = ghg_emissions,
                                 numberOfModelRuns = 10,
                                 functionSyntax = "plainNames")
GHG_simulation_scenarios$y[,"CO2eq_per_milk"]<-GHG_simulation_scenarios$y$pc_on_farm/GHG_simulation_scenarios$y$pc_milk_yield
dir.create("results")
save_temperature_scenarios(GHG_simulation_scenarios,"results","Baseline")

```

```{r, include=FALSE}
GHG_simulation_scenarios<-load_temperature_scenarios("results","Baseline")
class(GHG_simulation_scenarios)<-"mcSimulation"
```

Before we explore the results for the full population of farms, let's first look at one example. The following figure shows the greenhouse gas emissions, per farm and per head of cattle, for farm #5, which holds 12 head of cattle.

```{r, warning=FALSE, message=FALSE,fig.dim = c(8, 4), fig.cap = "***Figure 1.*** *Estimated total emissions from livestock (a) and emissions intensity of milk production (b) for a smallholder dairy farm in Kenya. Plausible estimates of both metrics are generated through Monte Carlo simulation, based on 1000 runs of a greenhouse gas emissions model, with uncertainty introduced by specifying inputs as probability distributions. The blue line indicates a precise estimate based on precise results from a farm survey and default parameters recommended by various guidance documents.*"}

scenarios<-read.csv("data/farm_scenarios.csv", fileEncoding = "UTF-8-BOM")
farm_5<-scenarios[,c(1,2,7)]

farm5_emissions<-scenario_mc(base_estimate = estimate_read_csv("data/livestock_ghg_input_table.csv"),
                             scenarios = farm_5,
                             model_function = ghg_emissions,
                             numberOfModelRuns = 1000,
                             functionSyntax = "plainNames")

precise_scenarios<-read.csv("data/farm_scenarios_precise.csv", fileEncoding = "UTF-8-BOM")

farm5_precise<-precise_scenarios[,c(1,2,7)]

farm5_emissions_precise<-scenario_mc(base_estimate = estimate_read_csv("data/livestock_ghg_input_table.csv"),
                             scenarios = farm5_precise,
                             model_function = ghg_emissions,
                             numberOfModelRuns = 10,
                             functionSyntax = "plainNames")

farm5a<-ggplot(data=farm5_emissions$y, aes(on_farm)) +
geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") +
  ylab("Relative probability") +
    xlab(expression(Total~emissions~from~livestock~(kg~CO[2]*'-eq'~year^-1))) + theme_bw() + ggtitle("a) Total emissions")

farm5_total <- farm5a + geom_vline(data=farm5_emissions_precise$y, aes(xintercept=on_farm),color="blue", lwd=3)

farm5_emissions$y[,"CO2eq_per_milk"]<-farm5_emissions$y$pc_on_farm/farm5_emissions$y$pc_milk_yield

farm5b<-ggplot(data=farm5_emissions$y, aes(CO2eq_per_milk)) +
geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") +
  ylab("Relative probability") +
    xlab(expression(Emissions~intensity~(kg~CO[2]*'-eq'~kg^-1))) + ggtitle("b) Emissions intensity")

farm5_emissions_precise$y[,"CO2eq_per_milk"]<-farm5_emissions_precise$y$pc_on_farm/farm5_emissions_precise$y$pc_milk_yield

farm5_intensity<-farm5b + geom_vline(data=farm5_emissions_precise$y, aes(xintercept=CO2eq_per_milk),color="blue", lwd=3) + theme_bw()

farm5_total+farm5_intensity

ggsave("figures/Fig_1_emissions_from_farm_5.png",width=9,height=5)

```

Now we look at the overall population.

```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 4), fig.cap = "***Figure 2.*** *Total greenhouse gas emissions from livestock (a) and emissions intensity of milk production (b) across 4140 simulated livestock-keeping households in Kenya (informed by data collected on 414 farms). Plausible estimates of both metrics are generated through Monte Carlo simulation, based on 1000 runs of a greenhouse gas emissions model, with uncertainty introduced by specifying inputs as probability distributions.*"}

all_farms_total<-ggplot(data=GHG_simulation_scenarios$y, aes(on_farm)) +
geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") +
  ylab("Relative probability") +
    xlab(expression(Total~emissions~from~livestock~(kg~CO[2]*'-eq'~year^-1))) + theme_bw() + ggtitle("a) Total emissions")

all_farms_intensity<-ggplot(data=GHG_simulation_scenarios$y, aes(CO2eq_per_milk)) +
geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") +
  ylab("Relative probability") + theme_bw() +
    xlab(expression(Emissions~intensity~(kg~CO[2]*'-eq'~kg^-1))) + ggtitle("b) Emissions intensity")

all_farms_total+all_farms_intensity

ggsave("figures/Fig_2_emissions_from_all_farms.png",width=9,height=5)
```

The

```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 4), fig.cap = "***Figure S1.*** *Greenhouse gas emissions from livestock per head of cattle across 4140 simulated livestock-keeping households in Kenya (informed by data collected on 414 farms). Plausible estimates are generated through Monte Carlo simulation, based on 1000 runs of a greenhouse gas emissions model, with uncertainty introduced by specifying inputs as probability distributions.*"}
all_farms_per_capita<-ggplot(data=GHG_simulation_scenarios$y, aes(pc_on_farm)) +
geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") +
  ylab("Relative probability") + theme_bw() +
    xlab(expression(Total~greenhouse~gas~emissions~per~capita~(kg~CO[2]*'-eq'~kg^-1))) + ggtitle("Emissions per capita")

all_farms_per_capita

ggsave("figures/Fig_S1_emissions_from_all_farms_per_capita.png",width=9,height=5)

```

We can now look at summary statistics of total and per capita emissions from cattle across livestock-keeping smallholder farms in Kenya. To exclude farms that do not appear to be specializing in livestock, some of which feature extremely high (up to infinity in cases where no milk is produced) emissions intensity, we also look at summary statistics across only the farms with per capita milk yields of >500 kg of milk.

```{r}

normal_milk_yield<-which(GHG_simulation_scenarios$y$pc_milk_yield>500)

dairy_households<-GHG_simulation_scenarios
dairy_households$x<-dairy_households$x[normal_milk_yield,]
dairy_households$y<-dairy_households$y[normal_milk_yield,]

emissions_summary<-rbind(data.frame("Emissions per farm"=round(quantile(GHG_simulation_scenarios$y$on_farm,
                                                                  c(0,0.05,0.5,0.95,1))/1000,1),
                                    "Emissions per capita"=round(quantile(GHG_simulation_scenarios$y$pc_on_farm,
                                                                    c(0,0.05,0.5,0.95,1))/1000,1),
                                    "Emissions intensity"=round(quantile(GHG_simulation_scenarios$y$CO2eq_per_milk,
                                                                   c(0,0.05,0.5,0.95,1)),1)),
                         data.frame("Emissions per farm"=round(quantile(dairy_households$y$on_farm,
                                                                  c(0,0.05,0.5,0.95,1))/1000,1),
                                    "Emissions per capita"=round(quantile(dairy_households$y$pc_on_farm,
                                                                    c(0,0.05,0.5,0.95,1))/1000,1),
                                    "Emissions intensity"=round(quantile(dairy_households$y$CO2eq_per_milk,
                                                                   c(0,0.05,0.5,0.95,1)),1)))
emissions_summary$Quantile<-c("Minimum","5% quantile", "Median", "95% quantile", "Maximum",
                              "Minimum","5% quantile", "Median", "95% quantile", "Maximum")
rownames(emissions_summary)<-NULL

kable(emissions_summary[,c(4,1,2,3)], booktabs = TRUE,
      col.names = c("Distribution attribute",
                    paste0("CO","$_2$","-eq farm","$^{-1}$"),
                    paste0("CO","$_2$","-eq head","$^{-1}$"),
                    paste0("CO","$_2$","-eq kg","$^{-1}$")),
      caption = "<p style='color:black'><i><b>Table S1.</b> Summary statistics of simulated greenhouse gas emissions from cattle across a population of cattle-keeping smallholder farms in Kenya</i></p>") %>% pack_rows(
  index = c("Emissions per farm" = 5, "Emissions per dairy-focused farm" = 5), background="lightgray") %>%
column_spec(2:4, width = "5cm") %>% row_spec(0, align = "c") %>% 
  add_header_above(c(" " = 1, "Emissions per farm" = 1, "Emissions per capita" = 1, "Emissions intensity" = 1))

```

\

We're interested in the suitability of total per capita milk production per year as an indicator of the greenhouse gas emissions intensity of milk production.

Here's the relationship between these two variables, for farms with per-capita milk yields of >500 kg per year.

```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 6), fig.cap="***Figure 3.***  *Relationship between milk yield per capita (x-axis) and emissions intensity of milk production for livestock-keeping smallholder households in Kenya, expressed as a population-level scatter plot (a) and an interpolated probability density surface (b). The vertical lines in (b) illustrate slices through the density surface at productivity levels of 2000 (red) and 3000 (blue) kg milk head^-1^ year^-1^ (the corresponding distributions are shown in Fig. 4b).*"}

plot1 <- ggplot(dairy_households$y, aes(x = pc_milk_yield, y = CO2eq_per_milk)) + 
  geom_point(shape = 20, color = "blue", size = 2) + 
  stat_smooth(method = "loess", fullrange = TRUE) +
  geom_density_2d() +
  geom_rug(color="lightskyblue3") + 
  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
       x=expression(Milk~yield~per~capita~(kg~head^{-1}~year^{-1})))+
  theme_bw() +
  theme(legend.position = c(0.15, 0.9))

dens1 <- ggplot(dairy_households$y, aes(x = pc_milk_yield)) + 
  geom_histogram(aes(y = ..density..),
                 colour = 1, fill = "white") +
  geom_density(fill="lightskyblue3") + 
  geom_histogram(aes(y = ..density..),
                 colour = 1, alpha=0) + 
  theme_void() + 
  theme(legend.position = "none") + ggtitle("a) Scatter plot of farm-scale data")

dens2 <- ggplot(dairy_households$y, aes(x = CO2eq_per_milk)) + 
  geom_histogram(aes(y = ..density..),
                 colour = 1, fill = "white") +
  geom_density(fill="lightskyblue3") + 
  geom_histogram(aes(y = ..density..),
                 colour = 1, alpha=0) + 
  theme_void() + 
  theme(legend.position = "none") + 
  coord_flip()

scatt<-dens1 + plot_spacer() + plot1 + dens2 + 
  plot_layout(ncol = 2, nrow = 2, widths = c(4, 1), heights = c(1, 4))

densi<-ggplot(dairy_households$y, aes(x = pc_milk_yield, y = CO2eq_per_milk)) +
  stat_density_2d(
    geom = "raster",
    aes(fill = after_stat(density)),
    contour = FALSE
    ) +
  scale_fill_viridis_c() + 
  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
       x=expression(Milk~yield~per~capita~(kg~head^{-1}~year^{-1})),
       fill="Probability \ndensity") +
  theme_bw() + geom_vline(xintercept=c(2000,3000),color=c("#F8766D","#00BFC4"),alpha=.5, lwd=2)

empty<-ggplot() +
    theme_void() +
    ggtitle("b) Interpolated probability density surface")

dens1 + plot_spacer() + empty + plot1 + dens2 + densi + 
  plot_layout(ncol = 3, nrow = 2, widths = c(4, 1), heights = c(1, 4))

ggsave("figures/Fig_3_scatter_and_surface_per_head_milk_emissions_intensity.png",width=10,height=6)
```

According to these distributions, the emissions intensity at a particular per capita milk yield can be described by a slice through the probability surface.

```{r varkernelslicerange, message=FALSE, warning=FALSE, fig.dim = c(12, 4), fig.cap="***Figure 4.*** *Greenhouse gas emissions intensity of milk production at specific levels of per-capita milk production among livestock-keeping smallholder farms in Kenya. Distributions are shown for (a) a production level of 2000 kg milk head^-1^ year^-1^, including a histogram of the random samples the distribution is based on, and (b) production levels of 2000 and 3000 kg milk head^-1^ year^-1^.*"}
slice_2000<-uncertainty::varslice_resample(in_var = dairy_households$y$pc_milk_yield,
                                 out_var = dairy_households$y$CO2eq_per_milk, 
                                 expectedin_var=2000,
                           n = 1000,
                           n_samples = 1000,
                           out_var_sampling = 1000)
slice_3000<-uncertainty::varslice_resample(in_var = dairy_households$y$pc_milk_yield,
                                 out_var = dairy_households$y$CO2eq_per_milk, 
                                 expectedin_var=3000,
                           n = 1000,
                           n_samples = 1000,
                           out_var_sampling = 1000)

one_slice<-ggplot(data=data.frame(CO2eq_per_milk=slice_2000$resampled), aes(CO2eq_per_milk)) +
geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") +
  ylab("Relative probability") +
    xlab(expression(Emissions~intensity~at~2000~kg~milk~per~capita~(kg~CO[2]*'-eq'~kg^-1))) + theme_bw() + ggtitle("a) Slice through the emissions intensity distribution")

slices<-rbind(data.frame(Values=slice_2000$resampled, Level="2000 kg"), data.frame(Values=slice_3000$resampled, Level="3000 kg"))

two_slices<-ggplot(data=slices, aes(Values, fill=Level)) +
 geom_density(alpha=.2) +
  ylab("Relative probability") +
    xlab(expression(Emissions~intensity~at~specific~milk~yield~per~capita~(kg~CO[2]*'-eq'~kg^-1))) + labs(fill='Productivity level') + theme_bw() + ggtitle("b) Two slices at different productivity levels")

one_slice + two_slices

ggsave("figures/Fig_4_slices.png",width=13,height=6)

```

By resampling from these distributions, we can determine the likelihood that a value of one of these distributions is greater than one of the other. We can systematically expand this to include more comparisons, in this case at intervals of 100 kg milk yield per capita.

```{r, fig.dim = c(8, 4), fig.cap="***Figure 5.*** *Probability of a particular farm B featuring lower emissions intensity of milk production than another farm A, based on observed per-capita milk production levels. Estimates are derived from pairwise comparisons of slices through a population-wide generalization surface of the relationship between per-capita milk production and emissions intensity. This generalization was generated through Monte Carlo simulation, based on scenarios representing observed production settings on 414 livestock-keeping smallholder farms in Kenya.*"}

milk_yields_to_sample<-seq(100,10000,100)
milk_yields_to_sample<-milk_yields_to_sample[which(milk_yields_to_sample<max(dairy_households$y$pc_milk_yield))]
milk_yields_to_sample<-milk_yields_to_sample[which(milk_yields_to_sample>min(dairy_households$y$pc_milk_yield))]
slices<-list()

# slice through the distribution at every 100 L milk yield per cow

for(i in 1:length(milk_yields_to_sample))
{slices[[i]]<-
  varslice_resample(dairy_households$y$pc_milk_yield,
                      dairy_households$y$CO2eq_per_milk,
                      expectedin_var = milk_yields_to_sample[i],
                      n_samples = 100000,
  out_var_sampling = 1000)$resample
 }

decreased_emissions_probability <-
  data.frame(Base_yield = NA,new_yield=NA,probability=NA)

for (i in 2:(length(milk_yields_to_sample) - 1))
{
  for (j in 1:length(milk_yields_to_sample))
    {decreased_emissions_probability[nrow(decreased_emissions_probability)+1,"Base_yield"]<-milk_yields_to_sample[i]
     decreased_emissions_probability$new_yield[nrow(decreased_emissions_probability)]<-milk_yields_to_sample[j]
     decreased_emissions_probability$probability[nrow(decreased_emissions_probability)] <-
      sum(slices[[i]] > slices[[j]]) / length(slices[[i]])
     decreased_emissions_probability$probability[nrow(decreased_emissions_probability)] <-
      sum(slices[[i]] > sample(slices[[j]], length(slices[[i]]))) / length(slices[[i]])

}
}

decreased_emissions_probability<-decreased_emissions_probability[which(!is.na(decreased_emissions_probability$Base_yield)),]

ggplot(data = decreased_emissions_probability, aes(Base_yield, new_yield, z =
                                                      probability)) + geom_contour_filled() +
  ggtitle("Probability of lower emissions intensity on farm B") + 
  labs(y=expression(Milk~yield~of~farm~B~(kg~head^{-1}~year^{-1})),
       x=expression(Milk~yield~of~farm~A~(kg~head^{-1}~year^{-1})),
       fill="Probability of \nlower emissions \nintensity on farm B") +
  stat_contour(breaks = c(0.9),col="black") + theme_bw()

ggsave("figures/Fig_5_probability_of_lower_emissions_intensity.png", width=7,height=5)

```

Now a description of the 5 interventions

```{r, warning=FALSE, fig.dim = c(8, 4), fig.cap="***Figure S2.*** *Observed distributions of milk yields for all cattle breeds encountered among dairy cows in a survey of livestock-keeping smallholder farms in Kenya.*"}
# breed intervention

cows<-Kenyaherds[which(Kenyaherds$animal_type %in% 4:6),]
cowcount<-data.frame(table(cows$breed))

breed_yield_plot<-ggplot(cows, aes(factor(breed),milk_yield)) +
  geom_violin(draw_quantiles=c(0.5),fill="light green") +
  xlab("Breed") + ylab(expression(Milk~yield~(kg~cow^{-1}~day^{-1}))) +
  geom_boxplot(width=0.1,fill="forestgreen") +
  theme_bw() +
  coord_flip() +
  scale_x_discrete(labels=c("Holstein Friesian (pure/improved)",
                            "Ayrshire (pure/improved)",
                            "Jersey (pure/improved)",
                            "Guernsey (pure/improved)",
                            "Cross-bred unknown",
                            "Boran",
                            "Local zebu")) +
  geom_text(data=cowcount ,aes(x = Var1, y = -2, label=paste("n =",Freq)),size = 4)

breed_milk_summary<-data.frame(Breed=c("Holstein Friesian (pure/improved)",
                            "Ayrshire (pure/improved)",
                            "Jersey (pure/improved)",
                            "Guernsey (pure/improved)",
                            "Cross-bred unknown",
                            "Boran",
                            "Local zebu"),
                            cow_number=cowcount$Freq,
                            milk_yield=round(ggplot_build(breed_yield_plot)$data[[2]]$middle,1),
                            intervention_factor= round(ggplot_build(breed_yield_plot)$data[[2]]$middle/
                             ggplot_build(breed_yield_plot)$data[[2]]$middle[1],2))
breed_milk_summary$intervention_factor[which(breed_milk_summary$Breed=="Boran")]<-1

breed_yield_plot

ggsave("figures/Fig_S2_breeds.png", width=9,height=4)

```
```{r}
kable(breed_milk_summary, col.names = c("Breed", "Number of dairy cows", "Daily milk yield (kg)", "Intervention factor"),
      caption = "<p style='color:black'><i><b>Table S2.</b> Mean milk yield across all dairy cow breeds encountered during a farm survey in Kenya</i></p>") %>%
column_spec(2:4, width = "5cm")

```

\

```{r}

milk_summary<-cbind(aggregate(cows$milk_yield, by=list(cows$breed), FUN=median),table(cows$breed))
milk_summary<-milk_summary[,c(1,4,2)]
colnames(milk_summary)<-c("Breed","n","Median_milk_yield")

milk_summary[,"Intervention_factor"]<-sapply(milk_summary$Median_milk_yield, function(x) min(x/milk_summary$Median_milk_yield[1],1))

live_weights<-aggregate(Kenyaherds$live_weight_LW, by=list(Kenyaherds$animal_type,Kenyaherds$breed), FUN=median)
colnames(live_weights)<-c("animal_type","breed","median_live_weight")
live_weights[,"Intervention_factor"]<-NA
for(i in 1:nrow(live_weights))
  live_weights$Intervention_factor[i]<-live_weights$median_live_weight[i]/
     live_weights$median_live_weight[which(live_weights$animal_type==live_weights$animal_type[i] &                                                                                                            live_weights$breed==1)]

# mature weight should probably be changed, and the code below does this. However, the mature weight is currently the same for all the breeds, so this accomplished nothing.

mature_weights<-aggregate(Kenyaherds$mature_weight_MW, by=list(Kenyaherds$animal_type,Kenyaherds$breed), FUN=median)
colnames(mature_weights)<-c("animal_type","breed","median_mature_weight")
mature_weights[,"Intervention_factor"]<-NA
for(i in 1:nrow(mature_weights))
  mature_weights$Intervention_factor[i]<-mature_weights$median_mature_weight[i]/
     mature_weights$median_mature_weight[which(mature_weights$animal_type==mature_weights$animal_type[i] &                                                                                                            mature_weights$breed==1)]

# same for the weight gain - doesn't depend on the breed at the moment, so all the factors are 1

weight_gains<-aggregate(Kenyaherds$weight_gain_WG, by=list(Kenyaherds$animal_type,Kenyaherds$breed), FUN=median)
colnames(weight_gains)<-c("animal_type","breed","median_weight_gain")
weight_gains[,"Intervention_factor"]<-NA
for(i in 1:nrow(weight_gains))
  weight_gains$Intervention_factor[i]<-weight_gains$median_weight_gain[i]/
     weight_gains$median_weight_gain[which(weight_gains$animal_type==weight_gains$animal_type[i] &                                                                                                            weight_gains$breed==1)]
weight_gains$Intervention_factor[which(weight_gains$median_weight_gain==0)]<-1

farms<-unique(Kenyaherds$QQNO)
breed_intervention<-data.frame(Farm=farms, milk_yield_4=1, milk_yield_5=1, milk_yield_6=1)

Kenyaherds_interventions<-Kenyaherds[,c("QQNO",
                                        "system",
                                        "animal_type",
                                        "breed",
                                        "live_weight_LW",
                                        "mature_weight_MW",
                                        "weight_gain_WG",
                                        "milk_yield",
                                        "dig_energy",
                                        "CP.")]

Kenyaherds_interventions[,"milk_change"]<-0
cows<-which(Kenyaherds_interventions$animal_type %in% 4:6)
Kenyaherds_interventions$milk_change[cows]<- sapply(cows,function(x) 1/milk_summary$Intervention_factor[milk_summary$Breed==Kenyaherds_interventions$breed[x]]  )

milk_change<-aggregate(Kenyaherds_interventions$milk_change[cows], by=list(Kenyaherds_interventions$QQNO[cows], Kenyaherds_interventions$animal_type[cows]), FUN=mean)

Milk_changes_4<-milk_change[which(milk_change$Group.2==4),]
breed_intervention$milk_yield_4[which(breed_intervention$Farm %in% Milk_changes_4[,1])]<-Milk_changes_4[,3]
Milk_changes_5<-milk_change[which(milk_change$Group.2==5),]
breed_intervention$milk_yield_5[which(breed_intervention$Farm %in% Milk_changes_5[,1])]<-Milk_changes_5[,3]
Milk_changes_6<-milk_change[which(milk_change$Group.2==6),]
breed_intervention$milk_yield_6[which(breed_intervention$Farm %in% Milk_changes_6[,1])]<-Milk_changes_6[,3]

for(i in 1:nrow(Kenyaherds_interventions))
  {Kenyaherds_interventions[i,"live_weight_change"]<-
    1/live_weights$Intervention_factor[which(live_weights$animal_type==Kenyaherds_interventions$animal_type[i]&
                                             live_weights$breed==Kenyaherds_interventions$breed[i])]
   Kenyaherds_interventions[i,"weight_gain_change"]<-
     1/weight_gains$Intervention_factor[which(weight_gains$animal_type==Kenyaherds_interventions$animal_type[i]&
                                              weight_gains$breed==Kenyaherds_interventions$breed[i])]
}

live_weight_agg<-aggregate(Kenyaherds_interventions$live_weight_change,
                           by=list(Kenyaherds_interventions$QQNO,Kenyaherds_interventions$animal_type), FUN=mean)
weight_gain_agg<-aggregate(Kenyaherds_interventions$weight_gain_change,
                           by=list(Kenyaherds_interventions$QQNO,Kenyaherds_interventions$animal_type), FUN=mean)

colnames(live_weight_agg)<-c("farm","animal_type","conv_factor")
colnames(weight_gain_agg)<-c("farm","animal_type","conv_factor")

for(i in 1:nrow(live_weight_agg))
  breed_intervention[which(breed_intervention$Farm==live_weight_agg$farm[i]),
                     paste0("live_weight_LW_",live_weight_agg$animal_type[i])]<-
    live_weight_agg$conv_factor[i]

for(i in 1:nrow(weight_gain_agg))
  breed_intervention[which(breed_intervention$Farm==weight_gain_agg$farm[i]),
                     paste0("weight_gain_WG_",weight_gain_agg$animal_type[i])]<-
    weight_gain_agg$conv_factor[i]

breed_intervention[is.na(breed_intervention)]<-1

x_breed_intervention <- GHG_simulation_scenarios$x

adoption_rate <- 1

intervention_adopters<-which(rbinom(nrow(x_breed_intervention),1,adoption_rate)==1)

for (i in 1:nrow(x_breed_intervention))
{
  if (i %in% intervention_adopters) # if farmers adopt, make changes in x file
    x_breed_intervention[i, colnames(breed_intervention)[2:ncol(breed_intervention)]] <-
      x_breed_intervention[i, colnames(breed_intervention)[2:ncol(breed_intervention)]] *
      breed_intervention[which(paste0("Farm_", breed_intervention$Farm) ==
                                 x_breed_intervention$Scenario[i]),
                         2:ncol(breed_intervention)]
}

breed_adopters<-which(!x_breed_intervention$milk_yield_4*x_breed_intervention$N_animal_type_4==GHG_simulation_scenarios$x$milk_yield_4*GHG_simulation_scenarios$x$N_animal_type_4 |
                        !x_breed_intervention$milk_yield_5*x_breed_intervention$N_animal_type_5==GHG_simulation_scenarios$x$milk_yield_5*GHG_simulation_scenarios$x$N_animal_type_5 |
                        !x_breed_intervention$milk_yield_6*x_breed_intervention$N_animal_type_6==GHG_simulation_scenarios$x$milk_yield_6*GHG_simulation_scenarios$x$N_animal_type_6)

## Herd intervention 1 - nobulls

x_nobull_intervention <- GHG_simulation_scenarios$x
nobull_adoption_rate <- 1.0
for (i in 1:nrow(x_nobull_intervention))
{
  if (rbinom(1, 1, nobull_adoption_rate)) # if farmers adopt, make changes in x file
    x_nobull_intervention$N_animal_type_1[i]<-0
}
nobull_adopters<-which(!x_nobull_intervention$N_animal_type_1==GHG_simulation_scenarios$x$N_animal_type_1)

## Herd intervention 2 - lessReplace

x_lessReplace_intervention <- GHG_simulation_scenarios$x
lessReplace_adoption_rate <- 0.23
for (i in 1:nrow(x_lessReplace_intervention))
{
  if (rbinom(1, 1, lessReplace_adoption_rate)) # if farmers adopt, make changes in x file
    x_lessReplace_intervention$N_animal_type_3[i]<-0
}
lessReplace_adopters<-which(!x_lessReplace_intervention$N_animal_type_3==GHG_simulation_scenarios$x$N_animal_type_3)

## Feed intervention 1 - Supplementary legume fodder

x_Calliandra_intervention <- GHG_simulation_scenarios$x
Calliandra_adoption_rate <- 0.3
for (i in 1:nrow(x_Calliandra_intervention))
{
  if (rbinom(1, 1, Calliandra_adoption_rate)) # if farmers adopt, make changes in x file
    {x_Calliandra_intervention$milk_yield_5[i]<-x_Calliandra_intervention$milk_yield_5[i]+0.156
     x_Calliandra_intervention$milk_yield_6[i]<-x_Calliandra_intervention$milk_yield_6[i]+0.156
     x_Calliandra_intervention$DE_perc[i]<-x_Calliandra_intervention$DE_perc[i]+0.14
     x_Calliandra_intervention$perc_crude_protein_diet[i]<-x_Calliandra_intervention$perc_crude_protein_diet[i]-0.07
     x_Calliandra_intervention$feed_prod_CO2[i]<-x_Calliandra_intervention$feed_prod_CO2[i]- (0.726 *
                                        sum(x_Calliandra_intervention[i,paste0("N_animal_type_",1:12)]))

  }
}
Calliandra_adopters<-which(!x_Calliandra_intervention$DE_perc==GHG_simulation_scenarios$x$DE_perc)

## Feed intervention 2 - Balanced diets

x_balance_intervention <- GHG_simulation_scenarios$x
balance_adoption_rate <- 0.14
for (i in 1:nrow(x_balance_intervention))
{
  if (rbinom(1, 1, balance_adoption_rate)) # if farmers adopt, make changes in x file
    {x_balance_intervention$milk_yield_5[i]<-x_balance_intervention$milk_yield_5[i]+0.75
     x_balance_intervention$milk_yield_6[i]<-x_balance_intervention$milk_yield_6[i]+0.75
     x_balance_intervention$feed_prod_CO2[i]<-x_balance_intervention$feed_prod_CO2[i]-(328.5 * 
                    sum(x_balance_intervention[i,paste0("N_animal_type_",1:12)]))
  }
}
balance_adopters<-which(!x_balance_intervention$feed_prod_CO2==GHG_simulation_scenarios$x$feed_prod_CO2)

```

To run the interventions, we need a function to run the model with a specific set of x values.

```{r}
# single model-run function
run_model<-function(x_data,model)
  {
  varnames<-colnames(x_data)[which(!colnames(x_data)=="Scenario")]
  model_function_ <- function(x) {

    e <- as.list(sapply(X = varnames, FUN = function(i) as.numeric(unlist(x[i]))))
    eval(expr = body(model), envir = e)
  }
  y <- do.call(what = rbind, args = lapply(X = apply(X = x_data, 
                                                     MARGIN = 1, FUN = model_function_), FUN = unlist))
  returnObject <- list(y = data.frame(y), x = data.frame(x_data))
  returnObject$call <- match.call()
  class(returnObject) <- cbind("mcSimulation", class(returnObject))
  return(returnObject)
}
```

```{r, eval=FALSE}
# run model for modified x file
breed_res <- run_model(x_breed_intervention, ghg_emissions)
nobull_res <- run_model(x_nobull_intervention, ghg_emissions)
lessReplace_res <- run_model(x_lessReplace_intervention, ghg_emissions)
Calliandra_res <- run_model(x_Calliandra_intervention, ghg_emissions)
balance_res <- run_model(x_balance_intervention, ghg_emissions)

breed_res$y[,"CO2eq_per_milk"]<-breed_res$y$pc_on_farm/breed_res$y$pc_milk_yield
nobull_res$y[,"CO2eq_per_milk"]<-nobull_res$y$pc_on_farm/nobull_res$y$pc_milk_yield
lessReplace_res$y[,"CO2eq_per_milk"]<-lessReplace_res$y$pc_on_farm/lessReplace_res$y$pc_milk_yield
Calliandra_res$y[,"CO2eq_per_milk"]<-Calliandra_res$y$pc_on_farm/Calliandra_res$y$pc_milk_yield
balance_res$y[,"CO2eq_per_milk"]<-balance_res$y$pc_on_farm/balance_res$y$pc_milk_yield

breed_res$call<-"multi-scenario call"
save_temperature_scenarios(breed_res,"results","Breed_intervention")
nobull_res$call<-"multi-scenario call"
save_temperature_scenarios(nobull_res,"results","Nobull_intervention")
lessReplace_res$call<-"multi-scenario call"
save_temperature_scenarios(lessReplace_res,"results","LessReplace_intervention")
Calliandra_res$call<-"multi-scenario call"
save_temperature_scenarios(Calliandra_res,"results","Calliandra_intervention")
balance_res$call<-"multi-scenario call"
save_temperature_scenarios(balance_res,"results","Balance_intervention")
```

```{r}
breed_res<-load_temperature_scenarios("results","Breed_intervention")
class(breed_res)<-"mcSimulation"
nobull_res<-load_temperature_scenarios("results","Nobull_intervention")
class(nobull_res)<-"mcSimulation"
lessReplace_res<-load_temperature_scenarios("results","LessReplace_intervention")
class(lessReplace_res)<-"mcSimulation"
Calliandra_res<-load_temperature_scenarios("results","Calliandra_intervention")
class(Calliandra_res)<-"mcSimulation"
balance_res<-load_temperature_scenarios("results","Balance_intervention")
class(balance_res)<-"mcSimulation"
```

```{r, fig.dim = c(8, 8), fig.cap="***Figure S2.*** *Intervention impact on per capita milk production and greenhouse gas emissions intensity.*"}

# make data.frame with all intervention results

interventions<-rbind(cbind(breed_res$y[breed_adopters,],
                           intervention="breed",
                           pc_milk_baseline=GHG_simulation_scenarios$y$pc_milk_yield[breed_adopters],
                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[breed_adopters],
                           Farm=breed_res$x$Scenario[breed_adopters]),
                     cbind(nobull_res$y[nobull_adopters,],
                           intervention="nobull",
                           pc_milk_baseline=GHG_simulation_scenarios$y$pc_milk_yield[nobull_adopters],
                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[nobull_adopters],
                           Farm=nobull_res$x$Scenario[nobull_adopters]),
                     cbind(lessReplace_res$y[lessReplace_adopters,],
                           intervention="lessReplace",
                           pc_milk_baseline=GHG_simulation_scenarios$y$pc_milk_yield[lessReplace_adopters],
                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[lessReplace_adopters],
                           Farm=lessReplace_res$x$Scenario[lessReplace_adopters]),
                     cbind(Calliandra_res$y[Calliandra_adopters,],
                           intervention="Calliandra",
                           pc_milk_baseline=GHG_simulation_scenarios$y$pc_milk_yield[Calliandra_adopters],
                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[Calliandra_adopters],
                           Farm=Calliandra_res$x$Scenario[Calliandra_adopters]),
                     cbind(balance_res$y[balance_adopters,],
                           intervention="balance",
                           pc_milk_baseline=GHG_simulation_scenarios$y$pc_milk_yield[balance_adopters],
                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[balance_adopters],
                           Farm=balance_res$x$Scenario[balance_adopters]))

interventions[,"milk_change"]<-interventions$pc_milk_yield-interventions$pc_milk_baseline
interventions[,"emissions_intensity_change"]<-interventions$CO2eq_per_milk-interventions$CO2_intensity_baseline
interventions[,"milk_change_relative"]<-(interventions$pc_milk_yield-interventions$pc_milk_baseline)/interventions$pc_milk_baseline*100
interventions[,"emissions_intensity_change_relative"]<-
  (interventions$CO2eq_per_milk-interventions$CO2_intensity_baseline)/interventions$CO2_intensity_baseline*100

interventions_normal_milk<-interventions[which(interventions$pc_milk_yield>500),]

intervention_change <- merge(data.frame(aggregate(interventions_normal_milk$emissions_intensity_change,
                                                  by = list(interventions_normal_milk$intervention),
                                                  FUN = mean)),
                             data.frame(aggregate(interventions_normal_milk$milk_change,
                                                  by = list(interventions_normal_milk$intervention),
                                                  FUN = mean)),
                             by = "Group.1")
colnames(intervention_change) <-
  c("intervention", "emissions_intensity_change", "milk_change")

intervention_colors<-palette.colors(n = 6, palette = "Okabe-Ito", alpha=1, recycle = FALSE)

interventions_scatter<-ggplot(data=interventions_normal_milk,
                        aes(x=milk_change,y=emissions_intensity_change)) +
  geom_point(aes(color=intervention),size=0.9) +
  theme_bw() +
  ggtitle("Intervention impact on milk yield and emissions") + 
  labs(y=expression(Change~'in'~emissions~intensity~(kg~CO[2]*'-eq'~kg^-1)),
       x="Change in per capita milk yield (kg)",
       color="Intervention") +
  scale_color_manual(values= c("breed" = intervention_colors[1],
                              "nobull" = intervention_colors[2],
                              "lessReplace" = intervention_colors[3],
                              "Calliandra" = intervention_colors[4],
                              "balance" = intervention_colors[6]),
                    labels = c("Breed upgrade", "No bulls", "Fewer males", "Calliandra feed", "Balanced diet")) +
  facet_wrap(~factor(intervention,
                     levels=c("breed","nobull","lessReplace","Calliandra","balance"),
                     labels=c("Breed upgrade", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet"))  ) +
  theme(legend.position = "none")

interventions_scatter

ggsave("figures/Fig_6a_intervention_impacts.png", width=9,height=6)

```

```{r, fig.dim = c(8, 8), fig.cap="***Figure 6.*** *Intervention impacts on relative per capita milk production and relative greenhouse gas emissions intensity, compared to a counterfactual farm with similar properties that does not implement the respective intervention.*"}

interventions_scatter_relative<-ggplot(data=interventions_normal_milk,
                        aes(x=milk_change_relative,y=emissions_intensity_change_relative)) +
  geom_point(aes(color=intervention),size=0.9) +
  theme_bw() +
  ggtitle("Intervention impact on milk yield and emissions") + 
  labs(y="Relative change in emissions intensity (%)",
       x="Relative change in per capita milk yield (%)",
       color="Intervention") +
  scale_color_manual(values= c("breed" = intervention_colors[1],
                              "nobull" = intervention_colors[2],
                              "lessReplace" = intervention_colors[3],
                              "Calliandra" = intervention_colors[4],
                              "balance" = intervention_colors[6]),
                    labels = c("Breed upgrade", "No bulls", "Fewer males", "Calliandra feed", "Balanced diet")) +
  facet_wrap(~factor(intervention,
                     levels=c("breed","nobull","lessReplace","Calliandra","balance"),
                     labels=c("Breed upgrade", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet"))  ) +
  theme(legend.position = "none")

interventions_scatter_relative

ggsave("figures/Fig_6_intervention_impacts_relative.png", width=9,height=6)

```

The herd structure interventions showed potential for causing major changes in milk yields and emissions intensity on particular farms. Plotting these impacts against the initial sets of relevant herd attributes serves to illustrate which farms are particularly responsive to these interventions (Fig. S3).

```{r, fig.dim = c(8, 4), fig.cap="***Figure S3.*** *Influence of the breed composition of farm-scale dairy cow populations on changes in milk yield and emissions intensity of milk production in response to a breed upgrade intervention*."}

interventions_diagnostics<-interventions

farm_structures<-data.frame(Farm_no=unique(Kenyaherds$QQNO))

for(f in farm_structures$Farm_no)
  {samp<-Kenyaherds[which(Kenyaherds$QQNO==f),]
   cows<-samp[which(samp$animal_type %in% 4:6),]
   farm_structures[which(farm_structures$Farm_no==f),"Farm"]<-paste0("Farm_",f)
   farm_structures[which(farm_structures$Farm_no==f),"breed_1"]<-length(which(cows$breed==1))/nrow(cows)
   farm_structures[which(farm_structures$Farm_no==f),"breed_2"]<-length(which(cows$breed==2))/nrow(cows)
   farm_structures[which(farm_structures$Farm_no==f),"breed_3"]<-length(which(cows$breed==3))/nrow(cows)
   farm_structures[which(farm_structures$Farm_no==f),"breed_4"]<-length(which(cows$breed==4))/nrow(cows)
   farm_structures[which(farm_structures$Farm_no==f),"breed_5"]<-length(which(cows$breed==5))/nrow(cows)
   farm_structures[which(farm_structures$Farm_no==f),"breed_6"]<-length(which(cows$breed==6))/nrow(cows)
   farm_structures[which(farm_structures$Farm_no==f),"breed_7"]<-length(which(cows$breed==7))/nrow(cows)
   farm_structures[which(farm_structures$Farm_no==f),"breed_8"]<-length(which(cows$breed==8))/nrow(cows)
   farm_structures[which(farm_structures$Farm_no==f),"unproductive_bulls"]<-
     length(which(samp$animal_type==1))/nrow(samp)
   farm_structures[which(farm_structures$Farm_no==f),"replacement_males"]<-
     length(which(samp$animal_type==3))/nrow(samp)

}

intervention_structures<-merge(interventions_diagnostics,farm_structures)
intervention_structures[,"not_unproductive_bulls"]<-1-intervention_structures$unproductive_bulls
intervention_structures[,"not_replacement_males"]<-1-intervention_structures$replacement_males

melted_breed<-melt(intervention_structures[,c("Farm",
                                              "intervention",
                                              "milk_change_relative",
                                              "emissions_intensity_change_relative",
                                              paste0("breed_",1:8),
                                              "unproductive_bulls",
                                              "replacement_males")],
                   id.vars=c("Farm",
                             "intervention",
                             "milk_change_relative",
                             "emissions_intensity_change_relative",
                             "unproductive_bulls",
                             "replacement_males"))
melted_breed<-melted_breed[which(!is.na(melted_breed$value)),]

#melted$Confidence_level_rounded<-round(melted$Confidence_level,2)*100

colnames(melted_breed)[which(colnames(melted_breed)=="value")]<-"breed_share"
colnames(melted_breed)[which(colnames(melted_breed)=="variable")]<-"breed"

metric_melt_breed<-melt(melted_breed,id.vars=c("Farm","intervention",
                                "unproductive_bulls","replacement_males","breed","breed_share"))
metric_melt_breed$value<-round(metric_melt_breed$value)
colnames(metric_melt_breed)[which(colnames(metric_melt_breed)=="value")]<-"change"
colnames(metric_melt_breed)[which(colnames(metric_melt_breed)=="variable")]<-"metric"

breedcol<-palette.colors(n = 8, palette = "Okabe-Ito", alpha=1, recycle = FALSE)

metrics <- c("Milk yield per capita", "Emissions intensity of milk production")
names(metrics) <- c("milk_change_relative","emissions_intensity_change_relative")

breed_diagnostic<-
  ggplot(metric_melt_breed[which(metric_melt_breed$intervention=="breed"),], aes(group=breed, fill=breed, y=breed_share, x=change)) + 
    geom_bar(position="fill", stat="identity") +
  scale_y_continuous(labels = scales::percent) + coord_flip() +
  labs(y="Cumulative share of breed in herd",
       x="Relative change in outcome metric (%)",
       fill="Breed of \ndairy cows") + scale_fill_manual(values= c("breed_1" = breedcol[1],
                                                                   "breed_2" = breedcol[2],
                                                                   "breed_3" = breedcol[3],
                                                                   "breed_4" = breedcol[4],
                                                                   "breed_5" = breedcol[5],
                                                                   "breed_6" = breedcol[6],
                                                                   "breed_7" = breedcol[7],
                                                                   "breed_8" = breedcol[8]),
                                               labels = c("Holstein Friesian (pure/improved)",
                                                          "Ayrshire (pure/improved)",
                                                          "Jersey (pure/improved)",
                                                          "Guernsey (pure/improved)",
                                                          "Cross-bred unknown",
                                                          "Sahiwal",
                                                          "Boran",
                                                          "Local zebu")) +
  theme_bw() + facet_wrap(~metric, 
  labeller = labeller(metric = metrics))

breed_diagnostic

ggsave("figures/Fig_S3_breed_diagnostics.png",width=8,height=5)

```

```{r, fig.dim = c(8, 4), fig.cap="***Figure S4.*** *Influence of the percentage of unproductive bulls in a cattle herd on changes in milk yield and emissions intensity of milk production in response to a 'retiring unproductive bulls' intervention*."}

melted_nobull<-melt(intervention_structures[which(intervention_structures$intervention=="nobull"),
                                            c("Farm",
                                              "milk_change_relative",
                                              "emissions_intensity_change_relative",
                                              "unproductive_bulls",
                                              "not_unproductive_bulls")],
                   id.vars=c("Farm",
                             "milk_change_relative",
                             "emissions_intensity_change_relative"))

colnames(melted_nobull)[which(colnames(melted_nobull)=="value")]<-"percentage"
colnames(melted_nobull)[which(colnames(melted_nobull)=="variable")]<-"animal_type"

metric_melt_nobull<-melt(melted_nobull,
                         id.vars=c("Farm",
                                   "animal_type",
                                   "percentage"))
metric_melt_nobull$value<-round(metric_melt_nobull$value)
colnames(metric_melt_nobull)[which(colnames(metric_melt_nobull)=="value")]<-"change"
colnames(metric_melt_nobull)[which(colnames(metric_melt_nobull)=="variable")]<-"metric"

metrics <- c("Milk yield per capita", "Emissions intensity of milk production")
names(metrics) <- c("milk_change_relative","emissions_intensity_change_relative")

metric_melt_nobull<-metric_melt_nobull[which(!is.na(metric_melt_nobull$change)),]

metric_melt_nobull$percentage<-round(metric_melt_nobull$percentage,2)

nobull_diagnostic<-
  ggplot(metric_melt_nobull, aes(group=animal_type, fill=animal_type, y=percentage, x=change)) + 
    geom_bar(position="fill", stat="identity") +
  scale_y_continuous(labels = scales::percent) + coord_flip() +
  labs(y="Cumulative share of breed in herd",
       x="Relative change in outcome metric (%)",
       fill="Animal type") + scale_fill_manual(values= c("unproductive_bulls" = "burlywood",
                                                         "not_unproductive_bulls" = "darkslategrey"),
                                               labels = c("Unproductive bulls",
                                                          "Other animals")) +
  theme_bw() + facet_wrap(~metric, 
  labeller = labeller(metric = metrics))

nobull_diagnostic

ggsave("figures/Fig_S4_nobull_diagnostics.png",width=8,height=5)

```

```{r, fig.dim = c(8, 4), fig.cap="***Figure S5.*** *Influence of the percentage of replacement males in a cattle herd on changes in milk yield and emissions intensity of milk production in response to a 'fewer replacement males' intervention*."}

melted_lessReplace<-melt(intervention_structures[which(intervention_structures$intervention=="lessReplace"),
                                            c("Farm",
                                              "milk_change_relative",
                                              "emissions_intensity_change_relative",
                                              "replacement_males",
                                              "not_replacement_males")],
                   id.vars=c("Farm",
                             "milk_change_relative",
                             "emissions_intensity_change_relative"))

colnames(melted_lessReplace)[which(colnames(melted_lessReplace)=="value")]<-"percentage"
colnames(melted_lessReplace)[which(colnames(melted_lessReplace)=="variable")]<-"animal_type"

metric_melt_lessReplace<-melt(melted_lessReplace,
                         id.vars=c("Farm",
                                   "animal_type",
                                   "percentage"))
metric_melt_lessReplace$value<-round(metric_melt_lessReplace$value)
colnames(metric_melt_lessReplace)[which(colnames(metric_melt_lessReplace)=="value")]<-"change"
colnames(metric_melt_lessReplace)[which(colnames(metric_melt_lessReplace)=="variable")]<-"metric"

metrics <- c("Milk yield per capita", "Emissions intensity of milk production")
names(metrics) <- c("milk_change_relative","emissions_intensity_change_relative")

metric_melt_lessReplace<-metric_melt_lessReplace[which(!is.na(metric_melt_lessReplace$change)),]

metric_melt_lessReplace$percentage<-round(metric_melt_lessReplace$percentage,2)

lessReplace_diagnostic<-
  ggplot(metric_melt_lessReplace, aes(group=animal_type, fill=animal_type, y=percentage, x=change)) + 
    geom_bar(position="fill", stat="identity") +
  scale_y_continuous(labels = scales::percent) + coord_flip() +
  labs(y="Cumulative share of breed in herd",
       x="Relative change in outcome metric (%)",
       fill="Animal type") + scale_fill_manual(values= c("replacement_males" = "burlywood",
                                                         "not_replacement_males" = "darkslategrey"),
                                               labels = c("Replacement males",
                                                          "Other animals")) +
  theme_bw() + facet_wrap(~metric, 
  labeller = labeller(metric = metrics))

lessReplace_diagnostic

ggsave("figures/Fig_S5_lessReplace_diagnostics.png",width=8,height=5)

```

Summary statistics of interventions for the counterfactual simulation

```{r}

ints<-c("breed","nobull","lessReplace","Calliandra","balance")

interventions_summary_milk<-data.frame(Intervention=c("Breed upgrade",
                                                 "Retiring unproductive bulls",
                                                 "Fewer replacement males",
                                                 "Calliandra feed",
                                                 "Balanced diet"))
interventions_summary_emissions<-interventions_summary_milk

for(i in 1:length(ints))
  {interventions_summary_milk[i,c("Weakest effect",
                                  "5% quantile",
                                  "Median effect",
                                  "95% quantile",
                                  "Strongest effect")]<-
    round(quantile(intervention_structures[which(intervention_structures$intervention==ints[i]),
                                           "milk_change_relative"],c(0,0.05,0.5,0.95,1),
                   na.rm=TRUE),1)
  interventions_summary_emissions[i,c("Weakest effect",
                                      "5% quantile",
                                      "Median effect",
                                      "95% quantile",
                                      "Strongest effect")]<-
    round(quantile(intervention_structures[which(intervention_structures$intervention==ints[i]),
                                           "emissions_intensity_change_relative"],c(1,0.95,0.5,0.05,0),
                   na.rm=TRUE),1)}

intervention_summary<-rbind(interventions_summary_milk,interventions_summary_emissions)

kable(intervention_summary, booktabs = TRUE,digits=0,
     # col.names = c("Intervention",
       #             paste0("CO","$_2$","-eq farm","$^{-1}$"),
        #            paste0("CO","$_2$","-eq head","$^{-1}$"),
         #           paste0("CO","$_2$","-eq kg","$^{-1}$")),
      caption = "<p style='color:black'><i><b>Table S3.</b> Summary statistics of the simulated impact of farm-scale interventions aimed at reducing greenhouse gas emissions from cattle on cattle-keeping smallholder farms in Kenya. Impacts are evaluated against a true counterfactual baseline, in which all farms are identical except for the simulated impacts of the interventions.  </i></p>") %>%
  pack_rows(index = c("Impact on milk yield per capita (%)" = 5, "Impact on emissions intensity (%)" = 5), background="lightgray") %>%
column_spec(2:6, width = "3cm")

```

\

These are noticeable changes, but can they be detected against the random variation that all farms experience? We'll now take another Monte Carlo sample, apply the same interventions to the same farms and evaluate the results again.

```{r, eval=FALSE}

## Run endline Monte Carlo simulation
GHG_simulation_scenarios_time2<-scenario_mc(base_estimate = estimate_read_csv("data/livestock_ghg_input_table.csv"),
                                            scenarios = read.csv("data/farm_scenarios.csv",
                                                                 fileEncoding = "UTF-8-BOM"),
                                            model_function = ghg_emissions,
                                            numberOfModelRuns = 10,
                                            functionSyntax = "plainNames")

GHG_simulation_scenarios_time2$call<-"multi-scenario call"
save_temperature_scenarios(GHG_simulation_scenarios_time2,"results","Endline_no_intervention")

```

```{r}

GHG_simulation_scenarios_time2<-load_temperature_scenarios("results","Endline_no_intervention")
class(GHG_simulation_scenarios_time2)<-"mcSimulation"

## Produce endline breed scenario
x_breed_intervention_endline <- GHG_simulation_scenarios_time2$x

for (i in breed_adopters)
{x_breed_intervention_endline[i, colnames(breed_intervention)[2:ncol(breed_intervention)]] <-
 x_breed_intervention_endline[i, colnames(breed_intervention)[2:ncol(breed_intervention)]] *
    breed_intervention[which(paste0("Farm_", breed_intervention$Farm) ==
                               x_breed_intervention_endline$Scenario[i]),
                       2:ncol(breed_intervention)]
}

## Herd intervention 1 - nobulls

x_nobull_intervention_endline <- GHG_simulation_scenarios_time2$x
x_nobull_intervention_endline$N_animal_type_1[nobull_adopters]<-0

## Herd intervention 2 - lessReplace

x_lessReplace_intervention_endline <- GHG_simulation_scenarios_time2$x
x_lessReplace_intervention_endline$N_animal_type_3[lessReplace_adopters]<-0

## Feed intervention 1 - Supplementary legume fodder

x_Calliandra_intervention_endline <- GHG_simulation_scenarios_time2$x
for (i in Calliandra_adopters)
{x_Calliandra_intervention_endline$milk_yield_5[i]<-x_Calliandra_intervention_endline$milk_yield_5[i]+0.156
 x_Calliandra_intervention_endline$milk_yield_6[i]<-x_Calliandra_intervention_endline$milk_yield_6[i]+0.156
 x_Calliandra_intervention_endline$DE_perc[i]<-x_Calliandra_intervention_endline$DE_perc[i]+0.14
 x_Calliandra_intervention_endline$perc_crude_protein_diet[i]<-x_Calliandra_intervention_endline$perc_crude_protein_diet[i]-0.07
 x_Calliandra_intervention_endline$feed_prod_CO2[i]<-x_Calliandra_intervention_endline$feed_prod_CO2[i]- (0.726 *
                                        sum(x_Calliandra_intervention_endline[i,paste0("N_animal_type_",1:12)]))
}

## Feed intervention 2 - Balanced diets

x_balance_intervention_endline <- GHG_simulation_scenarios_time2$x
for (i in balance_adopters)
{x_balance_intervention$milk_yield_5[i]<-x_balance_intervention$milk_yield_5[i]+0.75
 x_balance_intervention$milk_yield_6[i]<-x_balance_intervention$milk_yield_6[i]+0.75
 x_balance_intervention$feed_prod_CO2[i]<-x_balance_intervention$feed_prod_CO2[i]-(328.5 * 
                    sum(x_balance_intervention[i,paste0("N_animal_type_",1:12)]))
}

```

```{r, eval=FALSE}
# run model for modified x file (endline)
breed_end <- run_model(x_breed_intervention_endline, ghg_emissions)
nobull_end <- run_model(x_nobull_intervention_endline, ghg_emissions)
lessReplace_end <- run_model(x_lessReplace_intervention_endline, ghg_emissions)
Calliandra_end <- run_model(x_Calliandra_intervention_endline, ghg_emissions)
balance_end <- run_model(x_balance_intervention_endline, ghg_emissions)

breed_end$y[,"CO2eq_per_milk"]<-breed_end$y$pc_on_farm/breed_end$y$pc_milk_yield
nobull_end$y[,"CO2eq_per_milk"]<-nobull_end$y$pc_on_farm/nobull_end$y$pc_milk_yield
lessReplace_end$y[,"CO2eq_per_milk"]<-lessReplace_end$y$pc_on_farm/lessReplace_end$y$pc_milk_yield
Calliandra_end$y[,"CO2eq_per_milk"]<-Calliandra_end$y$pc_on_farm/Calliandra_end$y$pc_milk_yield
balance_end$y[,"CO2eq_per_milk"]<-balance_end$y$pc_on_farm/balance_end$y$pc_milk_yield

breed_end$call<-"multi-scenario call"
save_temperature_scenarios(breed_end,"results","End_Breed_intervention")
nobull_end$call<-"multi-scenario call"
save_temperature_scenarios(nobull_end,"results","End_Nobull_intervention")
lessReplace_end$call<-"multi-scenario call"
save_temperature_scenarios(lessReplace_end,"results","End_LessReplace_intervention")
Calliandra_end$call<-"multi-scenario call"
save_temperature_scenarios(Calliandra_end,"results","End_Calliandra_intervention")
balance_end$call<-"multi-scenario call"
save_temperature_scenarios(balance_end,"results","End_Balance_intervention")
```

```{r, include=FALSE}
breed_end<-load_temperature_scenarios("results","End_Breed_intervention")
class(breed_end)<-"mcSimulation"
nobull_end<-load_temperature_scenarios("results","End_Nobull_intervention")
class(nobull_end)<-"mcSimulation"
lessReplace_end<-load_temperature_scenarios("results","End_LessReplace_intervention")
class(lessReplace_end)<-"mcSimulation"
Calliandra_end<-load_temperature_scenarios("results","End_Calliandra_intervention")
class(Calliandra_end)<-"mcSimulation"
balance_end<-load_temperature_scenarios("results","End_Balance_intervention")
class(balance_end)<-"mcSimulation"
```

```{r, fig.dim = c(8, 8), fig.cap="***Figure S6.*** *Intervention impacts on milk production per capita and greenhouse gas emissions intensity, in comparison to random changes between two evaluation times. The gray dots illustrate random changes that occurred on farms that made no management changes, whereas the colored dots show changes on farms that implemented the respective interventions.*"}

# make data.frame with all intervention results

interventions_end<-rbind(cbind(breed_end$y[breed_adopters,],
                           intervention="breed",
                           pc_milk_baseline=GHG_simulation_scenarios$y$pc_milk_yield[breed_adopters],
                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[breed_adopters]),
                     cbind(nobull_end$y[nobull_adopters,],
                           intervention="nobull",
                           pc_milk_baseline=GHG_simulation_scenarios$y$pc_milk_yield[nobull_adopters],
                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[nobull_adopters]),
                     cbind(lessReplace_end$y[lessReplace_adopters,],
                           intervention="lessReplace",
                           pc_milk_baseline=GHG_simulation_scenarios$y$pc_milk_yield[lessReplace_adopters],
                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[lessReplace_adopters]),
                     cbind(Calliandra_end$y[Calliandra_adopters,],
                           intervention="Calliandra",
                           pc_milk_baseline=GHG_simulation_scenarios$y$pc_milk_yield[Calliandra_adopters],
                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[Calliandra_adopters]),
                     cbind(balance_end$y[balance_adopters,],
                           intervention="balance",
                           pc_milk_baseline=GHG_simulation_scenarios$y$pc_milk_yield[balance_adopters],
                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[balance_adopters]))

normal_end<-cbind(GHG_simulation_scenarios_time2$y,
                           pc_milk_baseline=GHG_simulation_scenarios$y$pc_milk_yield,
                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk)
normal_end[,"CO2eq_per_milk"]<-normal_end$pc_on_farm/normal_end$pc_milk_yield

interventions_end[,"milk_change"]<-interventions_end$pc_milk_yield-interventions_end$pc_milk_baseline
interventions_end[,"emissions_intensity_change"]<-interventions_end$CO2eq_per_milk-interventions_end$CO2_intensity_baseline
interventions_end[,"milk_change_relative"]<-(interventions_end$pc_milk_yield-interventions_end$pc_milk_baseline)/interventions_end$pc_milk_baseline*100
interventions_end[,"emissions_intensity_change_relative"]<-
  (interventions_end$CO2eq_per_milk-interventions_end$CO2_intensity_baseline)/interventions_end$CO2_intensity_baseline*100

normal_end[,"milk_change"]<-normal_end$pc_milk_yield-normal_end$pc_milk_baseline
normal_end[,"emissions_intensity_change"]<-normal_end$CO2eq_per_milk-normal_end$CO2_intensity_baseline
normal_end[,"milk_change_relative"]<-(normal_end$pc_milk_yield-normal_end$pc_milk_baseline)/normal_end$pc_milk_baseline*100
normal_end[,"emissions_intensity_change_relative"]<-
  (normal_end$CO2eq_per_milk-normal_end$CO2_intensity_baseline)/normal_end$CO2_intensity_baseline*100

interventions_normal_milk_end<-interventions_end[which(interventions_end$pc_milk_baseline>500),]
normal_normal_milk_end<-normal_end[which(normal_end$pc_milk_yield>500),]
normal_plot<-rbind(cbind(normal_normal_milk_end,intervention="breed"),
                   cbind(normal_normal_milk_end,intervention="nobull"),
                   cbind(normal_normal_milk_end,intervention="lessReplace"),
                   cbind(normal_normal_milk_end,intervention="Calliandra"),
                   cbind(normal_normal_milk_end,intervention="balance"))

intervention_colors<-palette.colors(n = 6, palette = "Okabe-Ito", alpha=1, recycle = FALSE)

interventions_scatter_end<-ggplot(data=normal_plot,
                        aes(x=milk_change,y=emissions_intensity_change)) +
  geom_point(color="gray",size=1) +
  geom_point(data=interventions_normal_milk_end,
                        aes(x=milk_change,y=emissions_intensity_change, color=intervention),size=0.8) +
  scale_color_manual(values= c("breed" = intervention_colors[1],
                              "nobull" = intervention_colors[2],
                              "lessReplace" = intervention_colors[3],
                              "Calliandra" = intervention_colors[4],
                              "balance" = intervention_colors[6]),
                    labels = c("Breed upgrade", "No bulls", "Fewer males", "Calliandra feed", "Balanced diet")) +
  theme_bw() +
  ggtitle("Intervention impact on milk yield and emissions") + 
  labs(y=expression(Change~'in'~emissions~intensity~(kg~CO[2]*'-eq'~kg^-1)),
       x="Change in per capita milk yield (kg)",
       color="Intervention")  +
  facet_wrap(~factor(intervention,
                     levels=c("breed","nobull","lessReplace","Calliandra","balance"),
                     labels=c("Breed upgrade", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet"))  ) +
  theme(legend.position = "none")

interventions_scatter_end

ggsave("figures/Fig_7a_intervention_impacts_endline.png", width=9,height=6)

```

```{r, fig.dim = c(8, 8), fig.cap="***Figure 7.*** *Relative intervention impacts on milk production per capita and greenhouse gas emissions intensity, in comparison to random changes between two evaluation times. The gray dots illustrate random changes that occurred on farms that made no management changes, whereas the colored dots show changes on farms that implemented the respective interventions.*"}

interventions_scatter_relative_end<-ggplot(data=normal_plot,
                        aes(x=milk_change_relative,y=emissions_intensity_change_relative)) +
  geom_point(color="gray",size=1) +
  geom_point(data=interventions_normal_milk_end,
                        aes(x=milk_change_relative,y=emissions_intensity_change_relative, color=intervention),size=0.8) +
  scale_color_manual(values= c("breed" = intervention_colors[1],
                              "nobull" = intervention_colors[2],
                              "lessReplace" = intervention_colors[3],
                              "Calliandra" = intervention_colors[4],
                              "balance" = intervention_colors[6]),
                    labels = c("Breed upgrade", "No bulls", "Fewer males", "Calliandra feed", "Balanced diet")) +
  theme_bw() +
  ggtitle("Intervention impact on milk yield and emissions") + 
  labs(y="Relative change in emissions intensity (%)",
       x="Relative change in per capita milk yield (%)",
       color="Intervention") +
  facet_wrap(~factor(intervention,
                     levels=c("breed","nobull","lessReplace","Calliandra","balance"),
                     labels=c("Breed upgrade", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet"))  ) +
  theme(legend.position = "none")

interventions_scatter_relative_end

ggsave("figures/Fig_7_intervention_impacts_endline_relative.png", width=9,height=6)

```

Summary statistics of interventions for the realistic-endline simulation

```{r}

ints<-c("breed","nobull","lessReplace","Calliandra","balance")

interventions_end_summary_milk<-data.frame(Intervention=c("Breed upgrade",
                                                 "Retiring unproductive bulls",
                                                 "Fewer replacement males",
                                                 "Calliandra feed",
                                                 "Balanced diet",
                                                 "No intervention"))
interventions_end_summary_emissions<-interventions_end_summary_milk

for(i in 1:length(ints))
  {interventions_end_summary_milk[i,c("Weakest effect",
                                  "5% quantile",
                                  "Median effect",
                                  "95% quantile",
                                  "Strongest effect")]<-
    round(quantile(interventions_normal_milk_end[which(interventions_normal_milk_end$intervention==ints[i]),
                                           "milk_change_relative"],c(0,0.05,0.5,0.95,1),
                                      na.rm=TRUE),1)
interventions_end_summary_emissions[i,c("Weakest effect",
                                      "5% quantile",
                                      "Median effect",
                                      "95% quantile",
                                      "Strongest effect")]<-
    round(quantile(interventions_normal_milk_end[which(interventions_normal_milk_end$intervention==ints[i]),
                                           "emissions_intensity_change_relative"],c(1,0.95,0.5,0.05,0),
                   na.rm=TRUE),1)}

interventions_end_summary_milk[6,c("Weakest effect",
                                      "5% quantile",
                                      "Median effect",
                                      "95% quantile",
                                      "Strongest effect")]<-
  round(quantile(normal_normal_milk_end[,"milk_change_relative"],
                 c(0,0.05,0.5,0.95,1),
                 na.rm=TRUE),1)
interventions_end_summary_emissions[6,c("Weakest effect",
                                      "5% quantile",
                                      "Median effect",
                                      "95% quantile",
                                      "Strongest effect")]<-
  round(quantile(normal_normal_milk_end[,"emissions_intensity_change_relative"],
                 c(1,0.95,0.5,0.05,0),
                 na.rm=TRUE),1)

intervention_end_summary<-rbind(interventions_end_summary_milk,interventions_end_summary_emissions)

kable(intervention_end_summary, booktabs = TRUE,digits=0,
      caption = "<p style='color:black'><i><b>Table S4.</b> Summary statistics of the simulated impact of farm-scale interventions aimed at reducing greenhouse gas emissions from cattle on cattle-keeping smallholder farms in Kenya. Results are based on comparison of a baseline scenario, where no farm was assumed to have implemented the interventions and an endline scenarios, where some farms adopted the mitigation measures. At both baseline and endline, variation in outcomes expected to result from random effects was simulated using Monte Carlo simulation (10 runs per farm, for 414 farms).</i></p>") %>%
  pack_rows(index = c("Impact on milk yield per capita (%)" = 6, "Impact on emissions intensity (%)" = 6), background="lightgray") %>%
column_spec(2:6, width = "3cm")

```

\

```{r, fig.dim = c(8, 4), fig.cap="***Figure 8. *** *Intervention impacts on milk yield and greenhouse gas emissions intensity at an endline sampling point vs. the initial sampling time. The colored background surface indicates the probability that increases in milk productivity are associated with decreasing emissions intensity.*"}

interventions_plot<-interventions_normal_milk_end
none_plot<-cbind(normal_normal_milk_end,intervention="none")
int_none_plot<-rbind(interventions_plot,none_plot)

interventions_probability<-ggplot(data = decreased_emissions_probability, aes(Base_yield, new_yield)) +
  geom_contour_filled(aes(z=probability)) +
  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
  labs(y=expression(Milk~yield~at~endline~(kg~head^{-1}~year^{-1})),
       x=expression(Milk~yield~at~baseline~(kg~head^{-1}~year^{-1})),
       fill="Probability of \nlower emissions \nintensity") +
  stat_contour(aes(z=probability),breaks = c(0.9),col="black") + theme_bw() +
  geom_point(data=int_none_plot,aes(x=pc_milk_baseline,y=pc_milk_yield), size=0.8, color="coral1") +
  ylim(c(0,NA)) + xlim(c(0,NA)) +
  facet_wrap(~factor(intervention,
                     levels=c("breed","nobull","lessReplace","Calliandra","balance","none"),
                     labels=c("Breed upgrade", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet", "No intervention"))  )

interventions_probability

ggsave("figures/Fig_8_probability_of_lower_emissions_intensity_interventions.png", width=7,height=5)

```

```{r, fig.dim = c(4, 6), fig.cap="***Figure 9.*** *Probability of Type I and Type II errors incurred by choosing specific levels of confidence that emissions intensity changes have actually occurred. Type I errors indicate the probability of failing to recognize households that implemented interventions, whereas the Type II error represents the percentage of households that did not make any changes, but are nevertheless rewarded due to random effects.*"}

# get closest probability value

for(i in 1:nrow(int_none_plot))
  {decreased_emissions_probability[,"diff"]<-
    abs(decreased_emissions_probability$Base_yield-int_none_plot$pc_milk_baseline [i]) +
    abs(decreased_emissions_probability$new_yield-int_none_plot$pc_milk_yield[i])
  int_none_plot[i,"confidence_level"]<-decreased_emissions_probability$probability[which(decreased_emissions_probability$diff==min(decreased_emissions_probability$diff))][1]
}

confidence_level_analysis<-data.frame(confidence=seq(0.25,1,0.01))

for(i in 1:nrow(confidence_level_analysis))
{# compute Type 1 errors for all interventions
  for(ints in c("breed", "nobull", "lessReplace", "Calliandra", "balance"))
    confidence_level_analysis[i,ints]<-
      length(which(int_none_plot$intervention==ints &
                     int_none_plot$confidence_level <
                     confidence_level_analysis$confidence[i])) /
      length(which(int_none_plot$intervention==ints)) * 100

  confidence_level_analysis$Type2error[i]<-
    length(which(int_none_plot$intervention=="none" &
                   int_none_plot$confidence_level>confidence_level_analysis$confidence[i])) /
    length(which(int_none_plot$intervention=="none")) * 100

}

confidence_level_analysis$confidence<-confidence_level_analysis$confidence*100
toplot<-melt(confidence_level_analysis,id.vars = "confidence")
toplot$variable <- factor(toplot$variable, levels = c("breed", "nobull", "lessReplace", "Calliandra", "balance", "Type2error"),
                  labels = c("Type I error - failure to reward 'Breed upgrade' adopters",
                             "Type I error - failure to reward 'Retire unproductive bulls' adopters",
                             "Type I error - failure to reward 'Fewer replacement males' adopters",
                             "Type I error - failure to reward 'Calliandra feed' adopters",
                             "Type I error - failure to reward 'Balanced diet' adopters",
                             "Type II error - rewarding non-adopters"))

ggplot(toplot, aes(x=confidence,y=value)) + geom_bar(stat="identity", fill="darkolivegreen") + facet_wrap(~ variable,nrow=6) + theme_bw() +
  labs(y="Error probability (%)", x="Required confidence that emission reductions have occurred (%)")

ggsave("figures/Fig_9_erorr_types.png", width=5,height=6)

```

and now for the sensitivity analysis

```{r, fig.dim = c(8, 5), fig.cap="***Figure 10.*** *Strength of the relationship of Monte Carlo model input variables with the emissions intensity of milk production, as indicated by the Variable-Importance-Score of a Partial Least Squares regression.*"}

pls_GHG<-dairy_households
pls_GHG$x<-pls_GHG$x[,which(!colnames(pls_GHG$x)=="Scenario")]

pls_out<-plsr.mcSimulation(object=pls_GHG, resultName = "CO2eq_per_milk")

legend_table<-read.csv("data/legend.csv", fileEncoding = "UTF-8-BOM")

plot_pls(pls_out,input_table=legend_table)
ggsave("figures/Fig_10_PLS_dairy_households.png", width=8,height=5)
```

# References